### 2017-05-03 papers/vph_2016/response_to_review.txt
The model as extracted from the abrg_local repository.

I'd like to thank the reviewer for their comments, which were
constructive and useful. I'll reply to the comments and indicate where
I have changed the manuscript below.

> Reviewer comments:

(Introduction)

> The introduction section depicts clearly the need of such a model.

(Materials & Methods)

> The second section presents a clear architecture of the platform
> components and the overall methodology. However, there is not a strong
> mathematical support of the algorithms used. The BRAHMS and SpineML
> could be just mentioned and not analysed as much, so that the gained
> space could be used for the mathematical background of the
> algorithms.

The description of BRAHMS and SpineML has been reduced. I've added the
equations governing the rate coded neural elements.

> More analysis should be performed regarding the connection
> of the Brain model and the muscular components, as it is there where
> the main novelty of the platform resides; thus, please explain
> thoroughly how the SC Map signal is converted into the 6
> time-dependent moto-neuron activation signals.

I have added additional detail to the description of the SC map -> SBG
conversion.

(Results)

> The results section contains a methodology portion regarding the
> model's training, so please consider of transferring it to the 2nd
> section.

This has been moved.

> Moreover, there is not any reference to the training data
> that were used, i.e. what exactly was the nature of the data input in
> the training process. Finally, a validation protocol of the simulated
> results is not mentioned at all, so please consider of adding any
> error metrics.

I have added a description of the data input for the training
process. I've not mentioned validation because the centroiding step,
ensures that only a single neural element drives the SC to SBG weight
maps and the range of inputs given ensures that every active element
in each weight map is trained.

> The reference section lacks of the titles of the journal papers and
> has to be revised.

I believe I have followed the short format for the bibliography, given
in the VPH2016 extended abstract template, which does not include
journal paper titles and so I have not made any changes here.