### 2017-12-08 papers/integration/omm_integ.tex
Minor fixes, following proof-read of manuscript

% While writing, don't stop for errors
\nonstopmode

% The Frontiers template.
\documentclass{frontiersSCNS}

\usepackage{url,hyperref,lineno,microtype}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage[onehalfspacing]{setspace}
\usepackage{gensymb} % For \degree

\linenumbers

% Trick to define a language alias and permit language = {en} in the .bib file.
% From: http://tex.stackexchange.com/questions/199254/babel-define-language-synonym
\usepackage{letltxmacro}
\LetLtxMacro{\ORIGselectlanguage}{\selectlanguage}
\makeatletter
\DeclareRobustCommand{\selectlanguage}[1]{%
  \@ifundefined{alias@\string#1}
    {\ORIGselectlanguage{#1}}
    {\begingroup\edef\x{\endgroup
       \noexpand\ORIGselectlanguage{\@nameuse{alias@#1}}}\x}%
}
\newcommand{\definelanguagealias}[2]{%
  \@namedef{alias@#1}{#2}%
}
\makeatother
\definelanguagealias{en}{english}
\definelanguagealias{eng}{english}
% End language alias trick

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{James {et~al.}}
\def\Authors{Sebastian James\,$^{1,2}$, Chris Papapavlou\,$^5$, Alexander Blenkinsop\,$^{1,2}$, \\
  Alexander Cope\,$^3$,  \\
  Sean Anderson\,$^{2,4}$, Konstantinos Moustakas\,$^5$, Kevin Gurney\,$^{1,2,*}$}

\def\Address{
$^{1}$Adaptive Behaviour Research Group, Department of Psychology,
  The University of Sheffield, Sheffield, UK \\
$^{2}$Insigneo Institute for in-silico Medicine,
  The University of Sheffield, Sheffield, UK \\
$^{3}$Department of Computer Science,
  The University of Sheffield, Sheffield, UK \\
$^{4}$Department of Automatic Control Systems Engineering,
  The University of Sheffield, Sheffield, UK \\
$^{5}$Department of Electrical and Computer Engineering,
  The University of Patras, Patras, Greece
}

\def\corrAuthor{Kevin Gurney}
\def\corrAddress{Department of Psychology, The University of Sheffield,
  Western Bank, Sheffield, S10 2TP, UK}
\def\corrEmail{k.gurney@sheffield.ac.uk}

\begin{document}

% Aliases
\newcommand{\ccg}{Cope-Chambers} % Omit Prescott-Gurney from the name,
                                % too unweildy.
\newcommand{\stob}{SpineML\_2\_BRAHMS}
% Emphasis and bold.
\newcommand{\e}{\emph}
\newcommand{\bol}{\textbf}
\newcommand{\blue}{\textcolor{blue}}
% I like my citations in blue
\newcommand{\mycite}[1]{\blue{\cite{#1}}}
% expon function is roman e
\newcommand{\expon}{\mathrm{e}}
% A font to indicate a branch of a github repository
\newcommand{\branch}[1]{\textbf{\texttt{#1}}}
\newcommand{\cmnt}[1]{\blue{#1}}
\newcommand{\dg}{$\degree$}

% Use colour to indicate changes made following review. Can be
% set to this 'black' colour prior to final publication:
\newcommand{\black}{\textcolor{black}}
% Colour for reviewer 1 changes:
\newcommand{\rvwcolour}{\textcolor{OrangeRed}}
% Colour for reviewer 2 changes:
\newcommand{\rvwcolourtwo}{\textcolor{Cerulean}}
% Commands to use in the text. Change 'rvwcolour/rvwcolourtwo' to
% 'black' to hide review changes.
\newcommand{\rvw}[1]{\rvwcolour{#1}}
\newcommand{\rvwt}[1]{\rvwcolourtwo{#1}}

\onecolumn
\firstpage{1}

\title[Integrated brain and biomechanics]{
Integrating brain and biomechanical models - a new paradigm
for understanding  neuro-muscular control
}

\author[\firstAuthorLast ]{\Authors}
\address{}
\correspondance{}
\extraAuth{} % If more than 1 corr. author refer to original template
             % and fill this in

\maketitle
\begin{abstract}

\section{}
To date, realistic models of how the central nervous system governs
behaviour have been restricted in scope to the brain, brainstem or
spinal column, as if these existed as disembodied organs.  Further,
the model is often exercised in relation to an in vivo physiological
experiment with input comprising an impulse, a periodic signal or
constant activation, and output as a pattern of neural activity in one
or more neural populations. Any link to behaviour is inferred only
indirectly via these activity patterns.
%
We argue that to discover the principles of operation of neural
systems, it is necessary to express their behaviour in terms of
physical movements of a realistic motor system, and to supply inputs
that mimic sensory experience. To do this with confidence, we must
connect our brain models to neuro-muscular models and provide relevant
visual and proprioceptive feedback signals, thereby closing the loop
of the simulation.
%
This paper describes an effort to develop just such an integrated
brain and biomechanical system using a number of pre-existing
models. It describes a model of the saccadic oculomotor system
incorporating a neuromuscular model of the eye and its six extraocular
muscles. The position of the eye determines how illumination of a
retinotopic input population projects information about the location
of a saccade target into the system.
%
A pre-existing saccadic burst generator model was incorporated into
the system, which generated motoneuron activity patterns suitable for
driving the biomechanical eye. The model was demonstrated to make
accurate saccades to a target luminance under a set of environmental
constraints.
%
Challenges encountered in the development of this model showed the
importance of this integrated modelling approach. Thus, we exposed
shortcomings in individual model components which were only apparent
when these were supplied with the more plausible inputs available in a
closed loop design. Consequently we were able to suggest missing
functionality which the system would require to reproduce more
realistic behaviour.
%
The construction of such closed-loop animal models constitutes a new
paradigm of \e{computational neurobehaviour} and promises a more
thoroughgoing approach to our understanding of the brain's function as
a controller for movement and behaviour.

\tiny
 \keyFont{
   \section{Keywords:}
   integrated brain biomechanics neuromuscular oculomotor saccade basal ganglia
 } % 5 to 8 keywords
\end{abstract}

\section{Introduction} \label{sec:introduction}

% Remove after review:
\bol{Note:} Changes based on Reviewer 1's comments \rvw{are in this colour}. For
Reviewer 2 \rvwt{we use this colour}.

% Background
The field of computational neuroscience has provided many \e{systems
models} of the
brain \citep{arai_two-dimensional_1994,gancarz_neural_1998,hazy_towards_2007,blenkinsop_frequency_2017}.
We refer to these as \e{mechanistic computational models}, meaning
models which consist of populations of neural elements, interconnected
in a biologically plausible manner, which simulate the operation of
the brain. Whilst they differ in scale and complexity, these models
all seek to describe the fundamental mechanisms behind common animal
behaviours such as locomotion, threat evasion, reaching or
feeding. However, none of the models cited here actually reproduce
these behaviours. In each case, the activity in a certain population
of neurons is taken to be representative of a behavioural outcome.
% Sometimes, ok to just have a brain model
In some cases, it \e{is} reasonable to take the activity of an
internal population within the brain model as being representative of
the induced behaviour. For example, a choice made in a \e{go/no-go}
task could be determined from activity in a population within a basal
ganglia model
\citep{nambu_discharge_1990,kuhn_event-related_2004}. The decision
to \e{go} is selected by a reduction of activity in this population;
maintenance of activity implies \e{no-go}. To validate the model, the
error rates which it generates could be compared with experimentally
determined error rates in primate subjects. We refer to this as
an \e{output assumption model} because the output is assumed to
signify behaviour. (An \e{input assumption model} assumes that sensory
input produces some particular form of neural activity in an input
population of the model.)

% Gap
However, we may be interested in reproducing accurate
simulated \e{trajectories}, in order to find out how degradation of
parts of the model affect movement. In Parkinson's Disease,
degradation of the dopamine neurons originating in the substantia
nigra pars compacta (SNc) causes
diskinesia \citep{galvan_pathophysiology_2008}, as well as abnormal
network activity in the basal
ganglia \citep{brown_dopamine_2001,mccarthy_striatal_2011}.  Sufferers
of the disease would be expected to produce abnormal
decision-making \e{and} movement trajectories in a
reach-to-the-correct-target task such as the one described in
\cite{james_target-distractor_2017}.
%
% Solution
A model which sought to explore in detail the effects of the SNc
degradation both on the decision making \e{and} on the movement
dynamics would need a physically accurate virtual arm, as well as
physically realistic sensory input for the brain. This is no less than
a complete model of those sections of the brain and body which act to
fulfil the task. Such a modelling effort, if successful, would result
in a virtual robot capable of expressing behaviour \e{in response to
sensory input from its environment}. This would represent a paradigm
shift in the field of computational neuroscience worthy of the new
name of \e{computational neurobehaviour}.

% Solution described means we now need to do background of eye movements:
In an attempt to build a model combining brain, realistic
biomechanics \e{and} sensory feedback, we sought to extend our
\rvw{previous work modelling the oculomotor system by adding a virtual,
biomechanical eye model able to make physically realistic
movements. The rotational state of the eye would then determine how
visual features in the virtual world were projected back into the
brain model.}
%
The existing
model \citep{cope_basal_2017} is already able to capture sensory input
and convert it into a neural signal, assumed to specify the target of
a \e{saccadic eye movement}; a fast movement of the eyes which directs
the fovea to a region of interest in the field of view.
%
The oculomotor system is an excellent candidate for modelling because
its movements can be specified with only three degrees of freedom,
making it one of the simplest neuro-muscular systems in the body.
%
It is nevertheless behaviourally interesting, as saccadic eye
movements reveal information about decision making at a subconscious
level
\citep{deubel_saccade_1996,reppert_modulation_2015,marcos_determining_2016}.
The modelling of the oculomotor system is served by a large body of
behavioural data describing
saccades \citep{tipper_reaching_2001,walker_effect_1997,casteau_effect_2012},
many anatomical studies of the neural substrates involved
\citep{meredith_intrinsic_1998,isa_intrinsic_2002,isa_exploring_2009}
and electrophysiological data linking these together
\citep{hepp_spatio-temporal_1983,dorris_neuronal_1997,mcpeek_competition_2003,vokoun_circuit_2011}.
Furthermore, in the context of building \e{behaving} systems, a
realistic mechanism for gathering visual information is a necessary
part of any model for which the behaviour requires visual attention
and decision making.  This is obvious from extrinsic
considerations---a subject must look at a scene to make decisions or
navigate within it. It also follows for \e{intrinsic} reasons. For
example, \cite{howard_hand_1997} showed that visual cues affect reach
trajectories and the same group later demonstrated that reaching
affects the saccadic system
\citep{tipper_reaching_2001} suggesting a close relationship between these
neural systems. \rvw{Building a behaving oculomotor system will
therefore assist future computational neurobehavioural modelling
efforts that involve reaching.}

Many neural populations are involved in the coding of saccadic eye
movements, only a very brief overview is given here; for a review,
see \cite{munoz_commentary:_2002}. One pathway takes information from
the retina directly into the superficial layers of the superior
colliculus in the brainstem
\citep{sterling_receptive_1971,linden_massive_1983,wu_involvement_1994}.
Activity within the superior colliculus then excites neurons in the
pons, medulla and rostral mid-brain \citep{sparks_brainstem_2002}
and \rvw{finally the motor neurons}, which innervate the extraocular
muscles
\citep{fuchs_firing_1970,sparks_brainstem_2002}. This direct pathway
is responsible for the low latency saccades called express saccades
\citep{schiller_effect_1987,edelman_activity_1996}.
%
Information from the retina is also processed by visual cortex which
feeds through to the frontal eye fields in which activity is related
to reflexive and voluntary
saccades \citep{schall_neural_1999}. Activity build-up in the frontal
eye fields is transferred to the intermediate layers of the superior
colliculus
\citep{stanton_frontal_1988-1} and is also processed by the basal ganglia,
which participates in the selection of the winning saccade end point
\citep{stanton_frontal_1988,hikosaka_role_2000}.
%
Although both cortical and subcortical paths produce a saccade target
signal in the superior colliculus, it is also possible for animals to
make relatively normal saccades even after the colliculus has been
ablated
\citep{wurtz_activity_1972,aizawa_reversible_1998},
though express saccades are lost with collicular lesions
\citep{schiller_effect_1987}. This makes the superior colliculus
a perplexing structure, being both critically involved in saccade
target specification \citep{sparks_sensory_1987} and saccade dynamic control
\citep{waitzman_superior_1991,goossens_optimal_2012} and yet dispensible.
The `backup pathway' likely incorporates the oculomotor vermis and
fastigial oculomotor region of the cerebellum which are known to
participate in the specification, dynamics and adaptation of saccadic
eye movements \citep{kleine_saccade-related_2003,takagi_effects_1998}.

There is a long history of modelling the oculomotor system. For a
comprehensive review, see \cite{girard_brainstem_2005}. Models of
individual \rvw{sub-systems} have been proposed for brainstem
\citep{robinson_oculomotor_1975,scudder_new_1988,gancarz_neural_1998},
cerebellum \citep{quaia_model_1999,dean_modelling_1995,dean_learning_1994}
and superior colliculus
\citep{massone_neural-network_1994,arai_two-dimensional_1994,moren_mechanism_2013,marino_spatial_2012}.
More recently, combined models have also been developed incorporating
sensory input \citep{cope_basal_2017} and driving \rvw{a second order
differential equation} representing the eye
\citep{tabareau_geometry_2007,nguyen_saccade_2014,thurat_biomimetic_2015}.
None of these models has yet fully closed the loop to produce a
behaving system operating freely within its environment. We argue that
developing integrated, closed-loop models of behaving systems offers
insights into the operation of neural systems that are not available
from input-~or output-assumption models.

\section{Material \& Methods} \label{sec:methods}

% A very short introduction to the section.
The integrated brain and biomechanical model described here is a
development of the model in \cite{cope_basal_2017}, referred to here
as the \e{\ccg~model}. This was a rate-coded neural network model
incorporating retinal populations, frontal eye fields (FEF), the basal
ganglia (BG), and the superior colliculus (SC).  The \ccg~model takes
as \e{input} the positions of luminances \rvwt{(of fixed shape and
intensity)} on a topographic map. \rvwt{While certain assumptions
were made about the input---that a luminant input excites activity on
a retinotopic layer, with computer code carrying out the
transformation achieved in the brain by a neural connectivity
map \citep{thivierge_topographic_2007}---it is nonetheless \e{not} an
input-assumption model according to our definition because the
activity generated in the neural input layer is modelled as a response
to the luminances, rather than being crafted.}  In the \ccg~model, the
centroid of the activity in the deep layers of superior colliculus was
assumed to accurately encode the location of the eye at the end of the
saccade
\citep{wurtz_activity_1972,robinson_eye_1972,van_gisbergen_collicular_1987,mcilwain_lateral_1982}.  This location was used to
recalculate the positions of the luminances in the eye's frame of
reference at each time step. \rvwt{Because a pattern of neural
activity in the output population was assumed to have a behavioural
outcome, it was thus an \e{output-assumption model}}.  The model
included no brainstem populations other than superior colliculus, nor
a neuromuscular model.

To the \ccg~model, we added a brainstem model and a biomechanical eye
model. \rvw{The rate-coded brainstem model was taken from the
literature} \citep{gancarz_neural_1998} as the best-of-breed saccadic
burst generator \citep{girard_brainstem_2005}.  The biomechanical eye
was implemented using the biomechanical modelling framework OpenSim;
the brain and brainstem were modelled using the SpineML toolchain.
These will be described below, \rvwt{along with a review of
the \ccg~model}, but first we will give a description of the
co-ordinate systems that were used.

\subsection{Co-ordinates in the world} \label{sec:methods:coords}

Before describing the biomechanical eye and the brain model, which
consisted of retinotopically mapped neural sheets, we describe the
co-ordinate system used in the world. The eye was located at the
origin of a three-dimensional, right-handed Cartesian co-ordinate
system, with its fovea directed in the $-z$ direction.  There was a
notional spherical screen which was also centred at the origin of the
co-ordinate system and had a radius of 50 (in arbitrary
units). The \e{fixation point} was the point on the screen at which
the eye was initially directed.  Onto the screen were projected target
luminances, each of which having a position described by two
co-ordinates; $\theta_{x}^{t}$, a rotation of the horizon plane about
the $x$ axis, and $\theta_{y}^{t}$, a rotation of the meridian plane
about the $y$ axis. The position is the intersection of these rotated
planes with the spherical screen (disregarding the intersection point
of these three surfaces behind the eye).
%
Note that a luminance with positive $\theta_{x}^{t}$ was above the
horizon of this world; one whose $\theta_{y}^{t}$ was positive lay to
the left of the world's meridian. For this reason, many of the figures
in this paper are plotted with $-\theta_{y}$ on the $x$-axis and
$\theta_{x}$ on the $y$-axis so that targets that lay up and to the
right in the world do so in the graphs, also.

Luminances were crosses of height and width subtending $\pm$3\dg~and
whose `bars' were 2\dg~thick. Luminances were oriented like $+$
symbols with their vertical bar aligned with the meridian plane and
their horizontal bar aligned with the horizon.

The eye's frame of reference was initially aligned with the world's
frame of reference.  At each timestep, the eye's rotational state
(described by the Euler rotations $\theta_x$, $\theta_y$, $\theta_z$)
was used to translate the three dimensional Cartesian co-ordinates of
the luminances in the world frame into co-ordinates in the eye
frame. The luminance co-ordinates in the eye's frame of reference were
used to determine the input to the brain model.

\subsection{Existing brain model} \label{sec:methods:existing}

% Describe what is modelled first. This para is 'The OM system
% performs action selection and the BG will do that for us.'
The brain model, excluding the brainstem, is a re-implementation of
the \ccg~model \rvwt{of reflexive saccadic behaviour \citep{cope_basal_2017}}.
\rvwt{Reflexive saccades are fast eye movements elicited by abrupt
changes in the peripheral visual scene (reflexive saccades can occur
also as a result of auditory and somatosensory stimuli, but these
modalities are ignored in this model). A reflexive saccade has a
starting position defined by the initial orientation of the eye and an
end-point position in which the eye is directed towards a new
target. Regardless of the number of targets within the visual scene,
the brain must choose one location as the end-point, because the eyes
can look only in one direction at a time. The functionality reproduced
by the \ccg~model is `the selection of the best target end-point for a
reflexive saccade'. A competition such as this between incompatible
movements is often referred to as an \e{action selection}
problem \citep{norman_attention_1986,maes_dynamics_1989,redgrave_basal_1999}.
The \ccg~model is therefore a model of action-selection in the
oculomotor system for reflexive saccades.  One hypothesis for the
r\^ole played by the basal ganglia (BG) is that the system
performs \e{action selection}
\citep{mink_basal_1996,redgrave_basal_1999,hikosaka_role_2000}.
The \ccg~model places the BG at the centre of the oculomotor system;
this follows the known anatomy of the
region \citep{hikosaka_role_2000} and provides a mechanism for action
selection of the best saccade. The BG receives input indirectly from
the superior colliculus, which has a retinotopic
arrangement \citep{ottes_visuomotor_1986}.}

% This para describes some anatomical stuff, and talks about what was
% omitted from the ccg model.
\rvwt{The BG receives excitatory inputs directly from retinotopic regions of
the cortex including the frontal eye fields (FEF), supplementary eye
fields (SEF), lateral intraparietal cortex (LIP) and dorsolateral
prefrontal cortex. The dorsolateral prefrontal cortex, which
participates in voluntary saccades
\citep{funahashi_prefrontal_1993,munoz_look_2004}, is not
modelled because the model concerns reflexive rather than voluntary
eye movements. Several other regions of the brain that are associated
with eye movements are also omitted from the model. The early visual
processing stream in cortex, from V1, through to the LIP is subsumed
into a `sustained retinal' signal which arrives at FEF. The
justification here is that the model reacts to simple luminant targets
and does not need to carry out the feature extraction performed by
these visual areas. The supplementary eye fields are involved in the
programming of saccade sequences \citep{tehovnik_eye_2000} and memory
guided
saccades \citep{chen_supplementary_1995,schlag_neurons_2002}. Lesions
of SEF do not affect visually guided
saccades \citep{gaymard_cortical_1998} and so the SEF is also omitted
from the model.}

% Now describe each bit a little more, with help of the CCG
% macroscopic architecture figure.
\rvwt{Fig.~\ref{fig:ccg_model}(a) shows the macroscopic architecture of
the \ccg~model. The figure shows the relationships between the retinal
input populations, the FEF, the populations comprising the BG
sub-system (the red border indicates that the box represents a number
of populations as a sub-system), the thalamus and the superior
colliculus. Excitatory connections are indicated with arrowheads;
inhibitory connections with circles in place of the arrowheads. The
blue and green connection lines indicate two thalamo-basal ganglia
loops, one cortical loop through FEF (green), the other a sub-cortical
loop through SC. It is important to note that although they are given
different colours in the diagram, these loops are in no way
independent, with loop activity combining both in thalamus and in the
basal ganglia and a direct excitatory, feed-forward connection from
FEF to SC.}

% More on BG model and mechanism; GPR recap
\rvwt{The basal ganglia sub-system is the most complex component of
the \ccg~model. The BG model is based on previous
work \citep{gurney_computational_2001,gurney_computational_2001-1} and
is referred to as the GPR model. The GPR model incorporates the
following main components of the primate
BG \citep{mink_basal_1996,wickens_basal_1997}: (i) The striatum (the
main input station to the BG) which is divided into two iterdigitated
populations of projection neurons expressing primarily D1 or D2-type
dopaminergic receptors (named Str\_D1 and Str\_D2); (ii) The
subthalamic nucleus (STN); (iii) the external segment of the globus
pallidus (GPe); (iv) the output nucleus relevant for saccadic
control---the substantia nigra pars reticulata
(SNr) \citep{hikosaka_role_2000}.}

% More GPR recap
\rvwt{The connectivity of the GPR model [Fig.~\ref{fig:ccg_model}(b)]
is constrained by the known anatomy and physiology of the
BG~\citep{bolam_synaptic_2000}. Physiologically, the only source of
glutamate within the BG is the STN, whose projections are therefore
excitatory; all other nuclei have GABAergic projection neurons and are
therefore inhibitory. The cortex sends glutamatergic projections to
both the Str\_D1 striatal population, which projects preferentially to
the SNr, and to Str\_D2, which projects primarily to
GPe \citep{gerfen_d1_1990}. The cortex also projects to the STN, which
sends diffuse projections to the SNr and
GPe \citep{parent_anatomical_1993}. The GPe projects to the SNr and
also projects back to the STN, completing a GPe--STN loop.}

% GPR model and action channels
\rvwt{The GPR model is arranged into `action channels';
Fig.~\ref{fig:ccg_model}(b) shows an example network containing three
channels. It is between these channels that competition occurs, with
the winning channel succeeding in reducing activity in the output
nucleus, SNr, and thereby disinhibiting its target. The complete
connectivity pattern for this small network is shown in
Fig.~\ref{fig:ccg_model}(b); the left channel in cortex innervates the
left channels of Str\_D1, STN and Str\_D2. Connections are one-to-one,
so it follows that the middle channel of cortex innervates the middle
channels of STN and the striatal populations and the right channel of
cortex innervates right channels in striatum and STN. Striatal
population channels also inhibit SNr and GPe on a one-to-one basis and
GPe feeds inhibition to SNr and STN in a one-to-one manner. The
outputs from STN however are not one-to-one. The output from all
channels of STN is summed together and then the sum is fed into each
channel of SNr and GPe. This models the diffuse excitation from STN
which has been observed in the BG \citep{parent_anatomical_1993}.}

% BG selection mechanism in GPR model
\rvwt{Within the BG, there are several mechanisms supporting
competitive processing for selecting channels whose inhibitory output
should be reduced. The selection mechanism of the GPR model is the
`off-centre, on-surround' scheme proposed
by \cite{mink_basal_1993}. The `on-surround' is provided by
diffuse, excitatory projections from the STN to the SNr. Focussed
inhibition from the Str\_D1 neurons in striatum contributes the
`off-centre' part of the mechanism. This arrangement leads to
selection behaviour via a release of target inhibition, since
channels that have strong salience (input) have weak output at the
level of SNr, and channels with weak salience have enhanced output.}

% BG mechanism - GPe's role
\rvwt{The GPe is not included in the centre-surround circuit described
above, but still plays a key r\^ole in selection. Operating alone, the
Str\_D1/STN/SNr circuit can suffer from the following problem: if the
input for all channels is relatively high, then the diffuse projection
from STN, which effectively supplies a sum of \e{all} of the STN
inputs to each channel in SNr, will provide so much excitation that
Str\_D1 may become unable to inhibit one of the channels in SNr and
selection may become
impossible. \cite{gurney_computational_2001,gurney_computational_2001-1}
showed that the inhibitory feedback from GPe to STN acts as an
`automatic gain control' to help prevent this from occurring.}

% Dopamine, tonic output.
\rvwt{At the neuronal level, the STN, GPe and SNr have tonic output
activity \citep{chevalier_disinhibition_1990,delong_primate_1985,kita_intracellular_1991}. This
is modelled using piecewise linear output functions with zero offsets,
$c$ (see Eq.~\ref{eq:lintransfunc}) but with noise added to the
input. In striatum, Str\_D1 and Str\_D2 have positive offset $c$,
mimicking the so-called `down-state' of medium spiny neurons which
have a resting potential far below spiking threshold and require
co-ordinated input to generate action
potentials \citep{wilson_origins_1996}.
% Note, in Alex's paper, there is an error with the offsets for D1,
% D2. These are set to -0.05 instead of 0.05. A quick trial with the
% model suggests this doesn't make a very large difference to the
% overall operation (you just see a background of noise around the
% signals in Str), and I have the sign correct in any case.
In addition, the Str\_D1 and Str\_D2 neurons are influenced by
dopamine in different ways; facilitating cortico-striatal transmission
at medium spiny neurons with D1
receptors \citep{hernandez-lopez_D1_1997,gonon_prolonged_1997} and
reducing transmission at those with D2
receptors \citep{delgado_inhibitory_1999}. These effects are modelled
using a dopamine parameter, $d$, which modulates the input activations
$a_{in}^{D1}$ and $a_{in}^{D2}$ as:
\begin{equation}
   a_{in}^{D1} = (0.2 + d)A
\end{equation}
\begin{equation}
   a_{in}^{D2} = (1 - d)A
\end{equation}
where $A$ is the input activation (see also
Eqns.~\ref{eq:a_in_d1} \& \ref{eq:a_in_d2}). For the `normal, healthy'
value for $d$ of 0.7, Str\_D1 activation is relatively enhanced
(0.9$A$); Str\_D2 activation is one third of this value (0.3$A$). The
major effect of this difference in the relative strength of the
activity in Str\_D1 versus Str\_D2 is simply that a change in the
level of activity in Str\_D1 affects the off-centre, on-surround
mechanism. The effect of varying the input into Str\_D2 is much more
subtle, with only a small change in the amount of inhibition fed from
GPe into STN (via a focussed, one-to-one connection) being affected by
the change, along with a small change in the inhibition fed into SNr
from GPe (also via a one-to-one connection). It is not possible to
ascribe the dynamic effect of the dopamine parameter to any single
population, because the activity is recurrently
connected through multiple loops. Thus, a line of reasoning such as
``reduced activity for a luminance in Str\_D2 will lead to less
inhibition in that region in GPe, which means that there will be
higher activity there, and hence more inhibition for that region
passed to STN leading us to expect less activity in STN'' is verified
by running a suitable simulation with the model, but the effect is
small. }
%
\rvwt{Note that
the effect of dopamine in the model is only to modulate the strength
of cortico-striatal synapses; no learning is modelled and so the
significance of dopamine as a prediction error signal is outside the
scope of the current work.}

% Change this - we already talked about channels in the context of the
% three channel model shown. Now talk about how we have a map of
% channels - a continuous space in the \ccg model. This can then lead
% on to Fig 3.
\rvwt{The GPR model in Fig.~\ref{fig:ccg_model}(b) has only three
channels, with the focussed inhibition from striatum to SNr and GPe
defined by a simple one-to-one scheme. The action channels represent
discrete, incompatible motor action choices. In the oculomotor model,
an action channel represents the end-point of saccade, and the
competition carried out in the basal ganglia is between potential
saccade end-points. However, eye movements have a \e{continuous}
end-point space; the eye can rotate to any orientation within its
biomechanically permissible range. Some end-points within this range
are mutually exclusive---it's not possible to look to the left and to
the right simultaneously---but \e{nearby} end-points are not
necessarily incompatible. A small enough error in the end-point of a
saccade will not prevent the eye from foveating on a target as the
foveal region of high visual acuity is not infinitesimally small. To
cope with this requirement, the populations within the oculomotor
basal ganglia are conceived of as two-dimensional topographic grids of
leaky integrator neural elements. Activity in each element corresponds
to a spatial location in the visual field. Neighbouring elements
correspond to locations which are close to each other in the visual
field. Focussed one-to-one projections in the GPR model are replaced
by projective fields with many weighted connections. Specifically,
each unit in Str\_D1 projects to a counterpart SNr$_j$ in SNr with
some weight $w_{max}$, but also connects to neighbouring nodes in SNr
with a weight given by $w_{max}.G(d)$, where $G(d)$ is a circularly
symmetric, 2D-Gaussian which is a function of distance $d$ from
SNr$_j$ [Fig.~\ref{fig:ccg_model}(c)]. A similar scheme applies for
the connectivity from Str\_D2 to GPe and for a number of the other
connections in the \ccg~model; in the SpineML implementation of the
model, this connectivity scheme is named `GaussianKernel'.
% NB - having this Fig. ref here is important for figure ordering!
Fig.~\ref{fig:brainmodel} shows a schematic of the SpineML implementation
of the model, based on a diagram produced by SpineCreator. Populations
for Str\_D1, Str\_D2, STN, SNr and GPe are
shown within the `Basal Ganglia' box. Input comes into the model via
the `World' population and the output population is SC\_deep. Compare
this diagram with Figs.~\ref{fig:ccg_model}(a) \&
(b). Fig.~\ref{fig:brainmodel} expands the `SC', `BG' and `slow retinal'
boxes from Fig.~\ref{fig:ccg_model}(a).}

% Now some FEF stuff taken from Alex's pre-pub
\rvwt{The frontal eye fields (FEF) are a key cortical area for the
generation of saccadic eye movements
\citep{hikosaka_role_2000,tehovnik_eye_2000,robinson_eye_1969,bruce_primate_1985}.
Saccadic targets are retinotopically mapped on its
surface \citep{robinson_eye_1969,bruce_primate_1985,sabes_parietal_2002},
and increased neural activity at a location on the map precedes a
saccade to that location. Importantly, the FEF is also associated with
visual decision making
\citep{thompson_visual_2005,schall_saccade_1995,monosov_measurements_2008,cohen_neural_2009}.
Thus, in a saccade choice, increased FEF activity is predictive of the
eye movement whether correct or
incorrect \citep{thompson_frontal_2005}, rather than of the correct
response.}

\rvwt{FEF neurons can be divided into three functional groups,
related to whether their activity corresponds with visual stimuli,
motor action, or both \citep{segraves_functional_1987}. The \ccg~model
simplifies this categorisation using a single layer of 50 by 50 units
representing the mean of all three groups. This layer therefore
responds to both visual stimuli and the buildup of activity associated
with motor (saccadic) action. The retina provides a persistent
luminance signal into the FEF through the dorsal visual
pathway \citep{ungerleider_two_1982} which is abbreviated in this
model to a direct connection with delay and noise.}

\rvwt{The FEF provides input into the
BG \citep{saint-cyr_organization_1990} (to Str\_D1, Str\_D2 and STN)
which, in turn, projects back to thalamus in a retinotopically
organised way \citep{middleton_basal_2000,lynch_input_1994}. In
addition, the thalamic targets of this path are regions with strong
reciprocal connections to the FEF \citep{mcfarland_thalamic_2002}. In
this way, the FEF forms channel-based loops through basal ganglia of
the kind described above. Such circuits formed the basis of the model
of \cite{humphries_role_2002}.
%In that model, the
%thalamic component included the thalamic reticular nucleus which
%implemented an additional selection mechanism (via a feed-forward
%competitive network within the cortico-thalamic loop). However, in the
%interests of model simplicity, this was not implemented in the \ccg
%oculomotor architecture.
The thalamo-cortical loop may be thought of as an integrator of
information, whose gain is modulated by inhibition from basal
ganglia \citep{chambers_mechanisms_2012,cope_biologically_2011}.}

\rvwt{The superior colliculus (SC) is a sub-cortical nucleus which
also plays a critical r\^ole in the generation of
saccades \citep{hikosaka_visual_1983}. Both FEF and SC have direct
connections to the saccadic burst generator (SBG, see
Sect.~\ref{sec:methods:brainstem}). If either is lesioned, the other
can direct gaze, following a period of
adjustment \citep{latto_effects_1977}, albeit with some persistent
deficits. The SC is also a direct target of output from the
SNr \citep{jayaraman_nigrotectal_1977,jiang_opposing_2003} and can be
influenced by the action selection mechanisms of the BG. In
particular, it forms a loop with BG, but unlike its cortical
counterpart in FEF, the input to basal ganglia comes via the thalamus
[Fig~\ref{fig:ccg_model}(a), blue arrows].}

\rvwt{While the SC has seven alternating cell-~and fibre-rich
layers \citep{wurtz_visual-motor_1980}, in most cases these are
categorised as the `superficial' and `deep' layers, which have
significantly different response properties. Cells in the superficial
layers, which receive input from the retina, are mainly visually
responsive, with a preferred response to phasic events (luminance
onsets and offsets) and movement on the visual
field \citep{goldberg_activity_1972}. In contrast, cells in the deep
layers receive multi-modal input, including inhibitory input from the
output structures of the BG \citep{jayaraman_nigrotectal_1977}, and
are directly involved in the generation of saccadic eye
movements. Saccade related activity in the deep layers appears to
generate saccades through `population coding', with a weighted sum of
activity across the retinotopy of SC determining the saccade target
\citep{lee_population_1988,van_opstal_role_1990,mays_dissociation_1980}.
The deep layers of SC receive input from the FEF in a topographic
manner \citep{stanton_frontal_1988,sommer_composition_2000}.}

\rvwt{The SC in the \ccg~model is based on the the model described
in \cite{arai_model_2005}, with the difference that the SNr input to
the SC is generated by the BG model, rather than being
hand-crafted. The SC model has a superficial and a deep layer, each of
which is a 2-D array of 50 by 50 leaky integrator units arranged in
the same retinotopic manner as the
FEF \citep{wurtz_visual-motor_1980}.}

% Foveal mask
\rvwt{The \ccg~model incorporates a special connectivity pattern for
visual input to the BG via cortical (FEF) and sub-cortical (thalamus)
pathways. Due to the retinotopic mapping (Sect.~\ref{sec:mapping}),
foveal luminances deliver a strong signal to the BG; roughly one third
of the map is activated for the foveal targets used in this work
(Fig.~\ref{fig:mapping}, red cross). This makes it virtually impossible
for a peripheral target (Fig.~\ref{fig:mapping}, yellow cross) to win
selection in the BG. Even if the peripheral target competed
successfully to generate a saccade, this process would cause a
significant delay, leading to latencies much larger than those
observed experimentally. To overcome this problem, the \ccg~model
incorporates a mechanism in which the synaptic strength of connections
between FEF, thalamus and striatum are reduced close to the fovea
according to a shifted hyperbolic tangent. This connection is named
`DecayingAtFovea' in the SpineML implementation and follows a modified
sigmoidal curve rather than tanh. In either case, the relation is
`S-shaped' and normalised to the range [0 1]. Far from the fovea
(where the S-shaped curve has the value $\approx$1), the connectivity
pattern looks almost identical to a one-to-one connection.}

% Retinal model
\rvwt{Input to the \ccg~model is provided through a simple retina
model which directly samples from a larger `world array' of pixel
values. In the current model, the input for the retina is named
`World' and is the retinotopic projection [Fig.~\ref{fig:brainmodel}(b)]
of the eye's field of view of the world [Fig.~\ref{fig:brainmodel}(a)]
and the luminant targets therein. The raw input in `World' is fed into
a population which adds noise, and then via a delayed connection to
FEF (the sustained retinal input path), to simulate processing through
the dorsal visual stream. It is also fed, without substantial delay,
into two leaky integrator layers (Retina\_1 and Retina\_2) with
different time constants, with the more slowly reacting layer
(Retina\_2) inhibiting its faster counterpart. The faster layer
responds quickly to the appearance of a prolonged stimulus before it
is inhibited by the slow layer, forming a phasic response to stimulus
onset. The mechanism ensures that phasic rather than tonic responses
arrive at the superficial SC from the retina.}

% CCG model output
\rvwt{The output of the \ccg~model is determined by the activity in the
SC\_deep population. The activity in SC\_deep is first transformed
from retinotopic co-ordinates into Cartesian co-ordinates. The
centroid of the activity is then computed. The position of this
centroid in the Cartesian frame determines the saccadic end-point. The
current model differs in that it does not compute a centroid, instead
feeding the SC\_deep activity into the saccadic burst generator.}

% Parameterisation here?
\rvwt{The \ccg~model was parameterised by tuning the model to perform
a prosaccade task in which a central luminance point was fixated by the
model. After a fixed duration, the fixation point was extinguished and
a target point of fixed luminance was presented. The model was tuned
so that the latency between the presentation of the target and the
initiation of an eye movement matched experimental
data \citep{reulen_latency_1984}, while also matching the
electrophysiological evidence of activity in a variety of brain
regions. The tuning of the BG model attempted to preserve as closely
as possible the weights used in the original paper
\citep{gurney_computational_2001}. Further details on
the parameterisation of the \ccg~model are given in \cite{cope_basal_2017}.}
% Differences between ccg and existing brain model or can I do that
% above? Better in the preceding text, I think and wherever it crops
% up.

\subsubsection{Components}

With the exceptions of the World and FEF\_add\_noise populations, each
neural element represents an activation; the activation is governed by
a first order differential equation specified in a \e{SpineML
component}. SpineML, which will be outlined in
Sect.~\ref{sec:methods:framework}, provides a means to mathematically
define the five distinct components in use in the brain model.

The \e{LINlinear} component, which is used in FEF, Thalamus, SC, SNr
and GPe populations, governs the activation $a$ with a first order
leaky integrator differential equation:
\begin{equation}
   \dot{a} = \frac{1}{\tau}(a_{in}-a)
\end{equation}
where $\tau$ is the time constant for the neural activation and
$a_{in}$ is the input to the neural element. $a_{in}$ is defined by an
activation input and a shunting inhibition input according to:
\begin{equation}
   a_{in} = A(1-s_a)+\alpha R_N
\end{equation}
Here, $A$ is the activation input and $s_a$ is the shunting inhibition
state variable whose value is related to the shunting input, $S$ by
\begin{equation}
   s_a = \begin{cases}
      S & S\leq 1 \\
      1 & S > 1
   \end{cases}
\end{equation}
$R_N$ is a random number drawn from a standard normal distribution
($\sigma$=1, $\mu$=0) and introduces noise to the activation of the
neural element, with the parameter $\alpha$ controlling the noise
amplitude.

The output, $y$, of LINlinear is related to the activation $a$ by the
piecewise linear transfer function
\begin{equation} \label{eq:lintransfunc}
   y(a) = \begin{cases}
      0   & a < c \\
      a-c & c \leq a \leq 1+c \\
      1   & a > 1+c
   \end{cases}
\end{equation}
where $c$ is a parameter defining the offset of the transfer
function. \rvwt{If $c<0$, then for zero activation ($a=0$), the output
will be positive. This simulates the effect of a neural population
having tonic firing. If $c>0$ then the output will be zero until the
activation exceeds $c$, simulating neurons which only fire when driven
by excitatory input.}

The \e{LINret} component used for the retinal populations is similar to
the LINlinear component, but with no intrinsic noise and no shunting
inhibitory input. It has a neural input which is identical to the
activation input $A$:
\begin{equation}
   a_{in} = A
\end{equation}
The \e{LINexp} component is a leaky integrator with an exponential
transfer function. It shares the same differential equation with
LINlinear, but has a different input equation and a different output
transfer function. It has the following equation for the neural
element input $a_{in}$:
\begin{equation}
   a_{in} = [A+N(a-V_{r}^{-})] (1-S) + 0.01 R_N
\end{equation}
where $A$ is the activation input and $N$ is an input which is
modulated by $V_{r}^{-}$, a reversal potential, and $a$, the current
activation of the element. These inputs are summed and then reduced by
a factor which is dependent on $S$, the shunting input. As in
LINlinear, $R_N$ introduces normally distributed noise to the element.

The output, $y$, of the LINexp component is given by
\begin{equation}
   y(a) = \begin{cases}
      \mathrm{e}^{a}-0.9   & \mathrm{e}^a \leq 1+0.9 \\
      1   & \mathrm{e}^a > 1+0.9
   \end{cases}
\end{equation}
This component is used in the subthalamic nucleus (STN) population, as
it gives a more physiologically accurate f-I
behaviour
\citep{wilson_model_2004,bevan_mechanisms_1999,hallworth_apamin-sensitive_2003}
which has been shown to allow the mapping of the basal ganglia network
architecture onto an optimal decision making
model \citep{bogacz_basal_2007}.

The \e{D1MSN} and \e{D2MSN} components are both leaky integrators, similar to
LINlinear. They differ in that they have no shunting inhibition. They
are used to model medium spiny neuron (MSN) populations in the
striatum. As they model the fact that most MSN neurons fall into two
groups; those expressing D1 dopamine receptors and those expressing D2
receptors, they have a dopamine parameter that modulates the input
activation, so that their equations for $a_{in}$ are thus:
\begin{equation} \label{eq:a_in_d1}
   a_{in}^{D1} = (0.2 + d)A + 0.01 R_N
\end{equation}
\begin{equation} \label{eq:a_in_d2}
   a_{in}^{D2} = (1 - d)A + 0.01 R_N
\end{equation}
where $d$ is the dopamine parameter. Varying dopamine from 0 to 1
enhances the activation in the D1 model, whereas it decreases the
activation of the D2 model elements, in line with experimental
observations \citep{harsing_influence_1997,gonon_prolonged_1997}.
Note that the equation for $a_{in}^{D1}$ differs from that used in
the \ccg~model, for which the cortico-striatal weights are multiplied
by $(1+d)$ rather than $(0.2+d)$. A typical value of $d$ is 0.7.

In the typical components given above, the value of the activation $A$
(and where relevant, the shunting input, $S$) is determined by summing
the weighted inputs to the population:
\begin{equation}
A = \sum_{i}w_i^{act} x_i^{act}
\end{equation}
\begin{equation}
S = \sum_{i}w_i^{sh} x_i^{sh}
\end{equation}
$w_i^{act}$ and $w_i^{sh}$ are, respectively, the weights of the
$i^{th}$ activation or shunting connections \rvwt{received by the component};
$x_i^{act}$ and $x_i^{sh}$ are the signals input to the activation and
shunting connections.

\subsubsection{Population activity and retinotopic mapping}
\label{sec:mapping}

Each population of 2500 neural elements was arranged in a 50 by 50
grid, with positions on the grid representing a retinotopic mapping
similar to that found empirically both in the superior
colliculus \citep{ottes_visuomotor_1986} and in visual cortex
\citep{eric_l._schwartz_computational_1980}
and assumed in this work to persist throughout the oculomotor system.

In a retinotopic mapping, the Cartesian co-ordinates of the
light-sensitive cells in the retina, whose density varies with
distance from the fovea, are transformed into the Cartesian
co-ordinates of the correspondingly active cells on the
colliculus. The mapping ensures that an even density of cells can be
maintained in the colliculus, but ensures that a group of adjoining,
active, retinal neurons will always activate an adjoining group of
neurons on the collicular surface.

The mapping turns out to resemble polar co-ordinates. That is, one
axis of the collicular surface specifies the eccentricity of a retinal
location (how far it is from the fovea) and the second axis specifies
the rotational angle of the retinal location; we therefore use the
convention of referring to the eccentricity axis on the colliculus as
$r$ and the rotation axis as $\phi$.

The \e{cortical magnification factor}, $M(r)$, gives the relationship
between the radial eccentricity $r$ and the retinal neural density. As
in
\cite{cope_basal_2017}, we use a first-order approximation of the form
for $M(r)$ given in \cite{rovamo_estimation_1979}:
\begin{equation} \label{eq:cmf}
M(r) = \frac{M_f}{1+\frac{r}{E_2}}
\end{equation}
The foveal magnification, $M_f$, is the magnification of the most
central region of the retina and has a value in the human of about 7.8
mm/\dg~\citep{rovamo_estimation_1979}.

In our model, $M_f$ is related to $W_{nfs}$, the width of the
retinotopic neural field, $W_{fov}$, the width of the eye's field of
view and $E_2$, the eccentricity at which the retinal density has
halved by:
\begin{equation} \label{eq:fm}
   M_f = \frac {W_{nfs}} {E_2\;\ln\left(\frac{W_{fov}}{2 E_2} + 1\right)}
\end{equation}
Here, $W_{nfs}$ is 50 (the side length of the 50x50 grid) and
$W_{fov}$ is set to 61\dg, a reduction from the biologically accurate
150\dg~due to the small number of neurons in the retinotopic neural
field. $E_2$ is
2.5 \citep{cope_basal_2017,slotnick_electrophysiological_2001}.

% Was looking for somewhere for this, possibly:
%The precision of the model to distinguish between
%targets, especially towards the periphery of the field of view was
%limited by the 50$\times$50 size of the neural layers. In order to optimise
%this precision, both the \ccg~model and our implementation of
%it had their effective eccentricity range limited. Our implementation
%was limited more than the Cope-Chambers model, because hills of
%activity in SC_deep had to be kept fully within each layer. When a
%hill `falls off the edge' of the layer, the amount of activation being
%passed to the SBG is reduced and the saccades become inaccurate. I
%didn't want to investigate effects at the edge of the SC, so I had an
%effective limit of the model of about 22 degrees of eccentricity for a
%saccade.

The mapping from the retinotopic co-ordinates in the brain to
rotational co-ordinates of the stimulus/response was written down by
\cite{schwartz_spatial_1977,eric_l._schwartz_computational_1980}
for measurements of striate cortex [visual stimulus to
electrophysiological response---\cite{daniel_representation_1961,
talbot_physiological_1941}] and by \cite{ottes_visuomotor_1986} for
superior colliculus data [electrophysiological SC stimulus to eye
movement response---\cite{robinson_eye_1972}].
% Ottes et al. fit to stimulus-to-movement data, and give a
% movement-to-collicular-coordinate mapping in the paper.
We used the following statement of this mapping to introduce stimuli
into the `World' input population of the brain model:
\begin{equation}\label{eq:thetas_to_phi}
   \phi = \frac{W_{nfs}} {2 \pi}\arctan\left(\frac{\theta_{y}^{t}}{\theta_{x}^{t}}\right)
\end{equation}
\begin{equation}\label{eq:thetas_to_r}
   r = M_fE_2\,\ln\left(\frac{1}{E_2}\sqrt{{\theta_{x}^{t}}^2 + {\theta_{y}^{t}}^2}+1\right)
\end{equation}
Note that we use $r$ and $\phi$ as the co-ordinates on the `collicular
surface'. Schwartz uses $r$ and $\phi$ as the polar coordinates of the
retinal stimulus; Ottes et al.~use $r$ and $\phi$ as polar coordinates
for the eye movement response; both use $u$ and $v$ as the Cartesian
co-ordinates of the neural map. We use $\theta_{x}^{t}$ and
$\theta_{y}^{t}$ to give Euler rotations for the retinal target
stimulus. Note also that the form of
Eqns.~\ref{eq:thetas_to_phi} \& \ref{eq:thetas_to_r} is slightly
different from that given in
\cite{ottes_visuomotor_1986} because our $\theta_{x}^{t}$ and
$\theta_{y}^{t}$ are not the polar co-ordinates used in that work.

The mapping encompases the entire visual field; the value of $\phi$ is
allowed to vary from 0\dg~to 360\dg~along its axis.  Effectively, the
two contralateral colliculi found in the biology are incorporated into
a single, square map, avoiding the need to carry out the kind of
`colliculus gluing' described in \cite{tabareau_geometry_2007}.

It is straightforward to show that the reverse mapping is given by:
\begin{equation}
   \theta_x = E_2 \left(e^{\frac{r}{M_f E_2}} - 1\right).\cos\left(\frac{2 \pi \phi}{W_{nfs}}\right)
\end{equation}
\begin{equation}
   \theta_y = E_2 \left(e^{\frac{r}{M_f E_2}} - 1\right).\sin\left(\frac{2 \pi \phi}{W_{nfs}}\right)
\end{equation}
where we have dropped the $t$ superscript on $\theta_x$ \& $\theta_y$,
as these equations transform a collicular location into rotations of
the eye.

Fig.~\ref{fig:mapping} shows the result of the mapping for a view of
two cross-shaped luminances. One cross illuminates the fovea, which
results in a large comb-shape of activity. The more peripheral cross
produces (in FEF) an indistinct object centred at a larger value of
$r$.

\subsubsection{Network}

Briefly, the model consists of input from the World population (see
Fig.~\ref{fig:brainmodel}, green population box) producing activity in an
`express' pathway to superior colliculus (purple) and simultaneously
in cortex, represented here by the FEF population (grey boxes in
Fig.~\ref{fig:brainmodel}). The express pathway causes short latency
activity in the superficial superior colliculus, which directly
innervates the deeper layers of the superior colliculus
(SC\_deep). Activity in FEF generates firing in a
thalamo-cortico-basal ganglia loop. The output of the basal ganglia is
the substantia nigra pars reticulata (SNr) which tonically inhibits
SC\_deep. If a location of activity in FEF is able to dominate
selection in the basal ganglia circuit, the corresponding location in
SNr will dis-inhibit and activity will build up in SC\_deep encoding
the saccade end point.

Connections shown in red are one-to-one connections; dark blue
projections indicate a connectivity pattern which `fans out' with a
2-D Gaussian kernel [Fig.~\ref{fig:ccg_model}(c)]; lighter blue
connections from the STN to SNr and GPe are diffuse, all-to-all
connections and projections coloured green are one-to-one connections
that decay towards the fovea so that foveal activity in FEF does not
swamp the basal ganglia which would prevent peripheral luminances from
ever being selected.
%
Note that SC\_deep contains two recurrent connections; one is
excitatory, with a Gaussian kernel mapping and the other implements
tecto-tectal inhibition, which increases the inhibition between
activity in opposite hemispheres of the field of
view \citep{gian_g._mascetti_tectotectal_1981,olivier_evidence_2000}
helping to resolve competition between saccades to the left and
right. The tecto-tectal inhibitory connection is \e{not} present in
the \ccg~model. In all other respects the model is as described in
\cite{cope_basal_2017}.
%
We have not listed the parameters of the network in tabular form here,
instead, the reader is referred to the SpineML declarative
specification of the model from the link given in SUPPLEMENTAL DATA.
The easiest way to access this information is by using SpineCreator.

% Section - description of new brainstem model
\subsection{Brainstem model} \label{sec:methods:brainstem}

We implemented a saccadic burst generator (SBG) based on the
connectivity outlined in \cite{gancarz_neural_1998}. The SBG network
for two of the model's six channels is shown in
Fig.~\ref{fig:sbg}. \rvw{In the brainstem model, we use the word `channel'
to mean a set of populations of neurons which are involved in
actuating a single extraocular muscle. SBG channels are arranged in
pairs, actuating opposing muscles. There is one pair of channels which
actuates the superior and inferior rectus muscles, causing vertical
rotations of the eye in a roughly parasaggital plane (the eye moves up
or down). Another pair actuates the lateral and medial rectus muscles,
causing horizontal rotations of the eye. The third pair actuates the
superior and inferior oblique muscles which contribute to vertical as
well as oblique rotations.} Activity from the output layer of superior
colliculus
%
(SC\_avg) is fed into each channel, which sums the activity it
receives and processes it in populations each of a single neural
element representing all the neurons in that population. Each channel
of the SBG functions to create the motor neuron activations that are
required to accelerate the eye in a particular direction, then hold
the eye in its new position against the returning force generated by
the elastic properties of the muscles.  The required motor neuron
activations are therefore a combination of features: a brief burst of
increased activity that accelerates the eye; followed by a period of
activity that is less than the burst firing rate but higher than the
tonic rate that exists when the eye is at the centre. This holds the
eye in its new position.

The SBG connectivity produces each of the these features separately,
then sums them to create the desired `bump and tonic' activation time
series.  The input to the first population in the SBG, the long-lead
burst neurons (LLBNs), is conceived as originating from one of the
deep layers of the superior colliculus.  The activity of the LLBNs are
passed to excitatory burst neurons (EBNs) which, in turn, inhibit the
LLBNs via the activity of the inhibitory burst neurons (IBNs).  This
feedback loop has a transmission delay, which allows activity to build
up in the EBNs before the inhibition is activated and the activity is
then reduced again.  This mechanism generates the `bump'.

The generation of the `tonic' phase of the required time series is
achieved simply by integrating the bump over time and multiplying by a
some small gain factor.  This is the function of the tonic neurons
(TNs).  The firing rate of the motor neuron defines the amount of
force applied to the eye by that muscle.  Thus, \rvw{the integral}
of the `bump'
defines how far the eye moves in that channel's direction.  The gain
and delay parameters in the LLBN-EBN-IBN-LLBN feedback loop therefore
have to be tuned such that the endpoint of the saccade is reasonably
accurate.  Furthermore the restoring force generated by the elasticity
of the muscles is dependent on the radial distance.  The value of the
new tonic firing rate, after the `bump' is dependent on the end
location of the eye.  If the ratio between the EBN firing rate and the
TN firing rate is not exactly correct, the eye will drift away from
the saccade endpoint after the saccade has been completed.  The EBN-TN
connection strength is therefore tuned such that the TN firing rate
yields a stable eye position across a range of eye eccentricities.

The omnipause neurons (OPNs) are tonically active and inhibit the
EBNs.  The activity of the OPNs is itself inhibited by activity in the
LLBNs.  The purpose of this arrangement is to ensure the eye does not
move in response to neural noise.

Each mean activity of all the neurons in each SBG population (except
the TNs) is defined by a single leaky integrator, first order
differential equation.
\begin{equation}\label{eq:LIN}
   \frac{da}{dt} = \frac{1}{\tau}(y-a)
\end{equation}
where $a$ is the activation of the nucleus, and $\tau$ is the time
constant of the nucleus. $y$ is a piecewise linear function of the
weighted sum of inputs to the nucleus and is given by
\begin{equation}
       y(IN) = \begin{cases}
       0    & IN \leq b \\
      IN-b   & b \leq IN \leq 1+b \\
      1   & IN \geq 1+b \\
   \end{cases}
\end{equation}
where $b$ is the $IN$ axis offset.  $IN$ is the weighted sum of inputs
to the nucleus and is given by,
\begin{equation}
    IN = \sum_{m}^{M} w_{mn} a_{m}
\end{equation}
where $a_{m}$ is the activation of the $m^{th}$ afferent nucleus.
$w_{mn}$ is the connection strength between the $m^{th}$ afferent
nucleus and the current nucleus. The activity of the TNs are defined
as
\begin{equation}
   \frac{da}{dt} = \frac{1}{\tau}\,y
\end{equation}
with the same piecewise linear transfer function as in the other SBG
populations.

% Section - description of biomechanical model
\subsection{Biomechanical eye} \label{sec:methods:eye}

The output signals of the brainstem's \rvwt{motoneuron (MN)
populations} are used to drive the biomechanical model. \rvwt{The MN
output signal in each brainstem channel is normalised in the range [0
1] and represents the mean firing rate of the neurons that innervate
the extraocular muscle for that channel.} The biomechanics are used
not only to get tangible feedback on the simulated saccades including
motion trajectories, but to add one more modelling dimension related
to the inertial properties of the eye plant including muscle
properties.

The biomechanical eye model, implemented using the OpenSim framework
\citep{seth_opensim:_2011}, is anatomically represented by a sphere of
uniform mass distribution. The diameter of the eye is 24 mm for
adults, with small variations between individuals; the mass of the eye
is 7.5 grams.  The eyeball is actuated by six extraocular muscles
(EOMs). The EOMs are arranged in three pairs forming a cone inside the
orbit with the apex being located inside the cranium in a tendonous
ring called the annulus of Zinn. An important feature of the
oculomotor system which greatly affects its overall behavior is the
existence of dynamic EOM pulleys. Their role is to guide the pivot
point of the EOMs. In our model, a pulley for each EOM has been
modeled by a point on the orbit whose location depends on the current
eye orientation.

% KM
\rvwt{An illustration of the biomechanical eye model is given in
Figure \ref{fig:bioeye}(a), while Figure \ref{fig:bioeye}(b) depicts
the head model used in the proposed framework.}

% KM
\rvwt{Two types of muscle models of different complexity are supported. The
first models muscles using linear path actuators. This simplistic
model of ideal muscles can be easily integrated with high level brain
models. As described above the muscles are wrapped around the eye.
The more complex model supported is based on the Thelen model
\citep{thelen_adjustment_2003} that is also supported by OpenSim and implements
Hill-type muscles. It includes realistic muscle wrapping geometric
entities of the muscle fibers, while it accommodates for both
activation and contraction dynamics.  The dynamics of muscular forces
can be split into: 1) The elasticity of the muscles. 2) A delay
between the onset of the afferent excitatory signal and the actual
muscle contraction, caused by the transmission time of the action
potentials and by the necessary calcium release at the muscle fibres.}

% EOM dynamics
The force applied by EOMs is controlled by an excitatory signal
supplied by motoneurons in the brainstem. The neural drive to produce
a saccadic eye movement can be characterized by a pulse component to
overcome the viscoelasticity of the orbital plant, a step component to
stabilize the eye in the new position, and a slide component that
models the gradual transition between the pulse and step.

% Orbital fat force
Passive forces due to the fatty tissues inside the eye orbit also
affect eye dynamics. Their role is critical in eliminating the
influence of head and body movements. We incorporated a custom torque,
$\mathbf{t}$, which acts like a rotational spring-damper apparatus,
resisting eyeball movements. It has elastic and viscous properties
governed by $\mathbf{t} = -K\mathbf{R}-C\mathbf{U}$ where $\mathbf{R}$
is the eye's orientation and $\mathbf{U}$ is its angular velocity. $K$
and $C$ are constants. A fuller description of the biomechanical model
can be found in \cite{papapavlou_physics-based_2014}.

% KM
\rvwt{Finally, numerical integration of the biomechanical eye model is based
on the Kutta-Merson integration method.}

\subsection{Model development framework} \label{sec:methods:framework}

The \ccg~model was originally developed to run on the BRAHMS model
execution framework
\citep{mitchinson_brahms:_2010,mitchinson_brahms_2015}. To run a
BRAHMS model, the researcher must develop \e{BRAHMS components} for
the various neural elements. A BRAHMS component is a programmatically
coded implementation of the behaviour of the component. It may have an
arbitrary number of inputs and outputs and may be written in C, C++,
Python or MATLAB. The \ccg~model components were hand written in C++
and MATLAB. A BRAHMS \e{SystemML} file describes how the different
components connect together and how data is passed between them
\citep{mitchinson_brahms:_2010}. The main BRAHMS program first
reads the SystemML file, then dynamically loads all the required
components before executing the system.

In the current work, the \ccg~model was reproduced using the
declarative SpineML markup
language \citep{alex_cope_spineml_2014,richmond_model_2014}, with the
help of the graphical SpineML model editing software called
SpineCreator
\citep{cope_spinecreator_2015,cope_spinecreator:_2016}. SpineML,
which is a development of the NineML specification
\citep{incf_task_force_on_multi-scale_modeling_network_2011},
describes neural populations and their projections in a highly
structured format in which neuron bodies, pre- and post-synapses are
described in terms of \e{SpineML components}. These are similar to the
components provided by BRAHMS, but in this case, the components are an
XML description of the functionality of the component, rather than a
programmatic implementation, with one XML file per component. A
SpineML \e{network layer} file then describes which components are
used in the model, and how they are connected together. Finally, a
number of SpineML \e{experiment layer} files specify how the model
described in the network layer can be executed. In the experiment
layer, the execution duration and timestep can be specified, along
with input conditions, connection lesions and component parameter
updates. A description of SpineML is given
in \cite{richmond_model_2014}; the definitive definition is found in
the schemas \citep{cope_spineml_2014}.  SpineCreator, in its r\^ole as
a graphical editor for the SpineML format, was used to generate the
SpineML files describing the model. It was also used to generate the
diagrams of the model.

As a declarative format for model specification, SpineML is agnostic
about how the model is executed. A number of simulation engines can be
utilised, including DAMSON \citep{richmond_damson_2015}, GeNN
\citep{nowotny_flexible_2011,nowotny_spineml_2014} and BRAHMS (used
here). The simulation engine incorporating BRAHMS is called \stob~
\citep{cope_spineml_2_brahms_2015}. \stob~is a collection of XSLT
stylesheets which first generate and compile C++ BRAHMS components
(which implement a simple, Forward-Euler solver) from the SpineML
component layer description files. \stob~then uses the SpineML
network and experiment layer files to generate a BRAHMS SystemML
description of the model. Finally, \stob~executes the model now
described entirely as a BRAHMS system, via a call to the BRAHMS
binary. A number of additional, hand-written components are present
in \stob~providing the inputs (constant inputs, time-varying inputs,
etc) which the modeller specifies in the experiment layer.

In addition to the brain model components, all of which are
code-generated using \stob~as described above, two hand-written
components are integrated into the model: The biomechanical eye model
and a sensory input component. The sensory input component takes the
eye's rotational state and the state of the experimental luminances
and projects a retinotopic activity map into the brain model. Both of
these BRAHMS components were hand-written in C++. To incorporate these
components into the SpineML model, a \stob~\e{external.xsl} file was
used.  The external.xsl file scheme for incorporating external BRAHMS
components into a SpineML model was a new \stob~feature motivated by
the current work.  Fig.~\ref{fig:modelframework} shows the workflow, in
which the model specification files (blue box - a combination of
SpineML files and C++ code), are processed (green box) into a BRAHMS
system (red box).

% Section - description of the integration. scientific and
% technical.
\subsection{Integrating the models and closing the loop}
\label{sec:methods:integrating}

The \ccg~model closes its loop by passing the centroid of activity in
SC\_deep (once it has surpassed a threshold) back to the code that
controls the world, which then uses this location to
instantaneously change the model's view of the world. In our extended
model, it was necessary to connect the output of the brain model back
to its input via the saccadic burst generator model and the
biomechanical eye. The resulting state of the eye, rather than the
centroid of the superior colliculus, was used to compute the input to
the brain, given the luminances visible in the world.

\rvwt{Thus, the information flow in the model is as follows:
Luminances in the world have their locations computed in the eye's
frame of reference, based on the rotational state of the eye. The
locations of the luminances on the retina are transformed into a
retinotopic co-ordinate system which determines the activity in the
`World' population (named to mean the `world as the brain sees it',
rather than the world frame of reference) which is the input for the
brain model. The target luminance for a saccade is selected, as
described earlier, via cortical and sub-cortical loops through the
basal ganglia model and activity for the winning end-point builds up
in the deep layer of superior colliculus. This activity excites the 6
channels of the saccadic burst generator in the correct proportions
for the saccade. The motoneurons, which are the output of the SBG,
send a rate-code signal (normalised between 0 and 1) into the
biomechanical eye model. The rotational state of the eye model is fed
back to participate in the computation of the retinotopic luminance
activity in `World', completing the loop.}

% Brain to SBG, seeing as that's how I explained it above.
A number of studies have considered the form of the connection between
the deeper layers of the superior colliculus and the saccadic burst
generator \citep{van_gisbergen_experimental_1985,ottes_visuomotor_1986,waitzman_superior_1991,groh_converting_2001,arai_two-dimensional_1994,goossens_dynamic_2006,tabareau_geometry_2007,van_opstal_linear_2008,goossens_optimal_2012},
which has become known as the spatial temporal transform (STT).  The
spatial aspect of the transform is thought to be implemented by a
weight-mapping \citep{tabareau_geometry_2007,arai_two-dimensional_1994}.
\rvwt{Although there is no definitive experimental proof for such a mapping,
there exists evidence for spatially variable synapse
density \citep{moschovakis_anatomical_1998,herrero_influence_1998} and
connection density \citep{grantyn_density_2002} and we therefore adopt
the idea.}
%
Arai and co-workers trained a 20 by 20 neural network model of the
superior colliculus to discover the weight map under the assumption of
2D Gaussian activation
profiles~\citep{arai_two-dimensional_1994}---\rvwt{that is, they
assumed that the activity in superior colliculus for any saccade was a
size-invariant 2D Gaussian hill of activity}.
%
The training approach of \cite{arai_two-dimensional_1994} was not
feasible in this study due to the length of time required to run our
model and its stochasticity, which meant multiple runs of the model
were necessary in order to generate output statistics.
%
\cite{tabareau_geometry_2007} wrote down a theoretical form of the weight
map, \rvwt{obtained by inverting the mapping} of \cite{ottes_visuomotor_1986}
and the assumption of invariant 2D Gaussian activity profiles in SC,
\rvwt{which is equivalent to:}
%
\begin{equation} \label{eq:tabareau_eq3}
w(r,\phi) = i\;e\,^{jr}\;\sin\left(l\phi + k\right)
\end{equation}
\rvwt{where $r$ and $\phi$ are co-ordinates on the collicular map and
$i$, $j$, $k$ and $l$ are parameters of the
function (compare with Eq.~3 of \cite{tabareau_geometry_2007}).}
%
As they found it closely resembles the results
of \cite{arai_two-dimensional_1994}, and it is a simple formulation,
we considered it as the means to generate the six weight maps in our
own model.
%
One barrier to the use of this weight map was the \ccg~model's
violation of the
\e{invariant integral hypothesis}. This states that the number of spikes
emitted by a neural element during a saccade (or in our model, the
integral of the neuron's output during the saccade) should be a
function only of its position within the hill of collicular
activity. That is, for any time-dependent hill of activity
$\mathcal{A}(\mathbf{z},t)$ at $\mathbf{z} = (r,\phi)$ on the collicular
surface, the integrated activity  $A_{\mathbf{x}}$ in an element at a
vector $\mathbf{x}$ away from $\mathbf{z}$ is
\begin{equation}
A_{\mathbf{x}} = \int_t \mathcal{A}(\mathbf{z}-\mathbf{x}, t)\;dt
\end{equation}
which is invariant for all $\mathbf{z}$. \rvwt{This requirement is
fulfilled by spatially invariant 2D Gaussian profiles, whose
time-course (how quickly they grow and then diminish) is always the
same.}

However, the very mapping on which the \cite{tabareau_geometry_2007}
result is based leads to a very \e{variant} activity profile \rvwt{in
the SC\_deep layer} of the \ccg~model. A luminance of a given size
which excites activity near to the fovea causes activity in a large
number of neurons in each retinotopic layer, whereas activity far from
the fovea excites a much smaller region. This effect is clearly
demonstrated in Fig.~\ref{fig:mapping} for equal sized targets both on
and distal from the fovea.

% Try to make this clearer for Reviewer 2, Methods Comments, point 5.
\rvwt{To understand the need for this invariance,
consider the effect of a 2D Gaussian hill in SC\_deep which elicits a
successful horizontal saccade of 10\dg. Activity from the 2D Gaussian,
passing through the weight maps will excite the superior and inferior
rectus channels by an equal, balanced amount, so these cancel out,
allowing the eye movement to be horizontal. The amount of activation
passed to the lateral rectus muscle results from a convolution of the
Gaussian and the exponential component of the weight map relationship
in Eq.~\ref{eq:tabareau_eq3}. If the Gaussian hill now appears further
along the collicular surface, coding for a 20\dg~saccade, \e{and also
becomes smaller}, we can still argue that the vertical component
signals to superior/inferior rectus muscles will cancel out, and we
could imagine that the exponential component of
Eq.~\ref{eq:tabareau_eq3} is correctly parameterised to compensate for
the smaller hill. However. Now consider a 2-D Gaussian hill which
codes for a 10\dg~saccade which is `up, and to the right' in equal
proportions. That means that the hill will sit on the boundary between
the weight maps for the `up', and the `right' muscles. Now, if the
hill moves to the $r=\mathrm{20}$\dg~location on the colliculus, \e{and also
reduces in size}, it will excite only the periphery of the sine; the
exponential increase of the map along $r$ is not guaranteed to
compensate for the reduction in the convolution of the Gaussian hill
and the sinusoidal component of the weight map.}
% Add figure for this.

This led us to hypothesise that the retinotopic mapping \rvwt{to the
SBG be preceded} by an associated widening projection field such that
the hill of activity in \rvwt{a `final' deep layer of superior
colliculus is invariant with position on the map}. There are a number of
locations in the system in which this widening projection field could
exist. It could be implemented in the projections between the retinal
populations and the superficial layer of SC along with the projection
between the World and the FEF population. However, this would affect
activity within the basal ganglia of the model, contradicting a result
in \cite{cope_basal_2017} which explains the `hockey stick' profile
for saccade latency as a function of saccade eccentricity. Instead, we
suggest that a widening projection field is encoded within the
superior colliculus itself, a complex, multi-layered structure which
could quite plausibly support such a function. Indeed, such widening
activity can be seen in the stimulation experiments
in \cite{vokoun_intralaminar_2010} and \cite{vokoun_response_2014}.
%
\rvw{In \cite{ghitani_excitatory_2014}, from the same research group,
evidence is presented for an excitatory and widely projecting pathway
from the stratum griseum intermediale (equivalent to our SC\_deep) to
the more superficial layers stratum opticum and stratum griseum
superficiale. Although this pathway is a `wide' projecting field, the
experiments do not indicate whether the projection \e{widens} along
the rostral-caudal axis of the SC. \cite{bayguinov_hard-wired_2015}
present evidence for another projecting field within SC whose
connectivity pattern \e{does} change along the rostral-caudal
axis. This projection is inhibitory in nature. Although neither of
these results precisely match the widening, excitatory projection
field hypothesised here, they do indicate that such connection
patterns are plausible.}
%
Although in this work we do not model
the SC in detail, we extended the model with a third functional layer
named SC\_deep2, shown in Fig.~\ref{fig:scdeep}(b) (\ccg~has only the two
layers SC\_sup and SC\_deep). We introduced a widening projection
based on a Gaussian projection field whose width, $\sigma(r)$ varies
in inverse proportion to the magnification factor, $M(r)$, given in
Eq.~\ref{eq:cmf} according to:
\begin{equation} \label{eq:sigmar}
\sigma(r) = \frac{m_{\sigma}}{M(r)} - \frac{m_{\sigma}}{M^0} + \sigma_0 \qquad r > r_0
\end{equation}
$m_{\sigma}$ is a scalar parameter which determines the `magnitude of
the widening'. $M^0$ is the `starting' magnification factor. Within
the foveal region ($0 \leq r \leq r_0$), the projection field is not
allowed to widen and so
\begin{equation} \label{eq:sigmar2}
\sigma(r) = \sigma_0 \qquad r \leq r_0
\end{equation}
which makes $\sigma_0$ the width of the Gaussian projection field
within the foveal region. (Note that the value chosen for the width of
the foveal region, $r_0$ is not identical to the foveal shift
parameter used in the \e{DecayingAtFovea} projections into striatum.)
The \e{Widening Gaussian} projection weight, $w(r,d)$ is then computed
as:
\begin{equation} \label{eq:widening}
w(r,d) = e^{-\frac{d^2}{2\sigma\left(r\right)^2}}
\end{equation}
where $d$ is the distance between the source and destination elements
in the collicular plane.  $m_\sigma$ was set to 50, $\sigma_0$ was
0.3, $M^0$ was 12.43 and $r_0$ was 20.
% Could have a figure here, made from widening_gaussian.m

A further issue regarding the use of the theoretical weight map
in \cite{tabareau_geometry_2007} was that it does not consider the
existence of the oblique extraocular muscles. There is evidence that
only two dimensional information is encoded in superior
colliculus
\citep{wurtz_activity_1972,hepp_monkey_1993,van_opstal_two-rather_1991},
but the eye is actuated by six extraocular muscles. In order to find
out a possible form for the input to the oblique muscles we carried
out a training process
% possibly descibe this training process in supplemental data
which depended on a centroid computation in SC\_deep and was designed
to maintain a null torsional eye rotation for all saccade
end-points. For the four rectus muscles, the resulting weight map
solutions resembled those found by
\cite{arai_two-dimensional_1994}. The trained
maps for the oblique muscles had a form very close to those for the
inferior and superior rectus channels, but with a smaller magnitude.
The inferior oblique map resembled the superior rectus map and the
superior oblique map resembled the inferior rectus. When parameterising
the theoretical weight maps, we set the inferior/superior
oblique maps to be 1/10$^{\mathrm{th}}$ of the superior/inferior rectus
maps, respectively. Interestingly, this suggests that there is a built-in
synergy between the vertical and oblique channels in the eye, although
the results will show there is some systematic change in the oblique error
with saccade end-point location.
% Could elaborate here on Listings Law, and why I didn't train to
% Listings Law. Do this in the chapter, rather than here in the paper.

\cite{tabareau_geometry_2007} give a formulation for the weight maps
in which it is possible to project both a positive and a negative
weight. In our model, all projections from SC\_deep are excitatory.
This means that each channel has a weight which follows the form:
\begin{equation} \label{eq:weightmaps}
w(r,\phi) = i\;e\,^{jr}\;\sin\left(\frac{2\pi\phi}{W_{nfs}} + k\right)
\end{equation}
where $i$, $j$ and $k$ are per-channel parameters for the weight
maps. $k$ is determined by the mapping. Only the positive part of the
sine is utilised. $i$ and $j$ are parameters to be found.

The saccadic burst generator model was originally conceived with the
assumption of a step input, which returns to zero activity at a
suitable time to curtail the saccade and avoid staircase
saccades \citep{gancarz_neural_1998}. In our model there is no such
mechanism to reduce activity in SC\_deep, and elsewhere.  Although a
successful, accurate saccade towards a target luminance will remove
the excitation which caused the activity in SC\_deep by bringing the
target luminance within the masked, foveal region, the activity in SC
decays too slowly to avoid additional saccadic movements. We found it
necessary to hypothesise an inhibitory feedback mechanism from the SBG
to the brain model.  This is shown in Fig.~\ref{fig:sbg}, which
indicates how the output from the inhibitory burst neurons (IBN) of
the SBG model are used to feed back an inhibitory signal to the
SC\_deep, thalamus and FEF populations in the brain model, resetting
them ready for the next saccade. \rvw{There is evidence for inhibitory
projections to SC from the propositus
hypoglossi nucleus \citep{corvisier_possible_1991}, which lies within the
brainstem, upstream from motoneurons, and has been shown to encode eye
velocity \citep{dale_nucleus_2013}.}

% SBG to worldDataMaker
The output signals from the six channels of the SBG were connected to
the six motoneuron inputs of the biomechanical eye. The signal was
normalised; a value of 1 meaning that all the motoneurons in the
output population were firing at their maximum rate and the force
exerted by the relevant extraocular muscle was maximal. Channels
innervated extraocular muscles as follows: Up: superior rectus; Down:
inferior rectus; Right: medial rectus; Left: lateral rectus; Z+:
superior oblique; Z-: inferior oblique. Because the medial rectus
induces a rightward rotation of the eye, our single virtual eye is
a \e{left} eye. The OpenSim implementation of the biomechanical eye
was `wrapped' (in the software sense) in a BRAHMS component. This made
it possible to integrate the OpenSim model into the BRAHMS
framework. The wrapper ensured that the input and output signals were
correctly transferred and, importantly, handled the disparity in the
solver timesteps used in the OpenSim model (25~ms) and the neural
model (1~ms). This was achieved by having the BRAHMS wrapper create a
separate thread to run the OpenSim model. The BRAHMS wrapper component
was called on each 1 ms timestep, receiving the instantaneous
activations from the motoneurons in the SBG. These activations, and
the current simulation time, were written into a shared memory area,
accessible by the OpenSim thread. Running independently, the OpenSim
thread would update its inputs (using the most recent values in the
shared memory area) whenever the simulation time had increased by 25
ms. It would then recompute its outputs (the rotational state of the
eye) and write these into the same shared memory. The BRAHMS wrapper
would update its outputs whenever they were changed in the shared
memory by the OpenSim thread. A direct connection of the six outputs
of the BRAHMS eye model component to the six inputs of the
worldDataMaker BRAHMS component was specified in
the \stob~external.xsl file.

% worldDataMaker to brain:
The eye model outputs its rotational state at each timestep. The
rotational state is used to compute the view of the world in the eye's
frame of reference. To simplify the calculation, the luminances exist
on a spherical surface at the centre of which is the eye. A hand-coded
BRAHMS component called worldDataMaker computes the projection of the
luminances into the eye's frame of reference and then converts this
representation into a retinotopic map to pass into the brain
model. The input to the brain model is thus able to change
continuously, on every timestep, rather than in a step-wise fashion
when a saccade occurs, as in the \ccg~model.

In the worldDataMaker BRAHMS component, the rotational state of the
eye was used to construct Euler rotation matrices which transformed
between the world's frame of reference and the eye's frame of
reference. The worldDataMaker component received a specification of
the world luminances in a JSON file called luminances.json at the
start of each simulation. This file specified the position,
shape, size, luminance, appearance time and disappearance time of an
arbitrary number of luminances. With this information, the
instantaneous rotational state of the eye and the parameters of the
retinotopic transform, it was able to compute the instantaneous input
to the brain model.

The final models, on which the results of this paper are based are
named `TModel3', `TModel4' and `TModel5'. Descriptions of these, and
earlier versions of the model can be found in the code repository
given in SUPPLEMENTAL DATA.

\section{Results} \label{sec:results}

\subsection{Weight maps} \label{sec:results:weightmaps}

We found the best parameters for the exponential in
Eq.~\ref{eq:weightmaps} ($i$ and $j$) by a manual tuning
process. After selecting values for $i$ and $j$ in either the
horizontal or vertical/oblique channels, we ran the model 6 times at
each of 8 target eccentricities (7\dg--14\dg) which were purely in the
direction of the newly parameterised channel. The training saccades
were produced as described below in Sect.~\ref{sec:results:singlesaccades},
with the same fixation and target luminances (crosses of magnitude 0.2
and 0.3) but with the fixation offset and target onset occurring at
0.2~s. We measured the end-point of the saccade by detecting the
location at which the saccade velocity had dropped below 0.005 of its
peak. We iterated until the mean saccade endpoint plotted versus
target was close to the ideal straight line---see
Fig.~\ref{fig:saccvstarg}(a) \& \rvw{(b)}. We applied the same parameters to
both directions of each channel; $i_{up} = i_{down} = 0.00195$,
$j_{up} = j_{down} = 0.075$, $i_{left} = i_{right} = 0.0016$ and
$j_{left} = j_{right} = 0.067$.

The resulting weight maps (where the oblique maps are
1/10$^{\mathrm{th}}$ of the vertical maps, as described earlier) are
shown in Fig.~\ref{fig:weightmaps}.
%
\rvw{First, recall that the $r$ axis of the neural surface corresponds
to the amplitude of a saccade and the $\phi$ axis indicates the polar
direction of the saccade, as described in Sect.~\ref{sec:mapping} and
Fig.~\ref{fig:mapping}. Fig.~\ref{fig:weightmaps}(a) shows the weight
map for the muscle which rotates the eye to the left. As we modelled a
left eye, this actuates the lateral rectus muscle. The exponential
rise of Eq.~\ref{eq:weightmaps} [for experimental evidence,
see Figs.~7 \& 8 of \cite{herrero_influence_1998}] is seen in the
$r$ direction; as $r$ increases, so the connection strength to the SBG
channel rises exponentially. The connection strength is greatest along
the centre line, for a value of $\phi$ which corresponds to a purely
leftward movement. Note that $\phi$ is presented in neural
coordinates, and not in degrees or radians; $\mathrm{1}\leq\phi\leq\mathrm{50}$
corresponds to a range of 0\dg~to 360\dg; $\phi=\mathrm{38.5}$ corresponds to
movements left. The connection strength drops away sinusoidally as
$\phi$ moves away from the centre line at $\phi=\mathrm{38.5}$. In regions of
the map for which there is no leftward movement, that is, in the half
of the map which corresponds to any movement with a rightward
component, the `left' weight map is 0. Fig.~\ref{fig:weightmaps}(d)
shows the weight map for rightward movements, actuating the medial
rectus muscle of the eye. The line of maximum connection strength is
along $\phi=\mathrm{13.5}$. The map is a mirror of
Fig.~\ref{fig:weightmaps}(a), reflected about the line $\phi\mathrm{26}$. The
`left' and `right' weight maps are orthogonal; the non-zero region of
the `left' map is zero in the `right' map and vice
versa. Fig~\ref{fig:weightmaps}(b) \& (d) show the weight maps for
downward and upward eye movements; the `down' map activates the SBG
channel for the inferior rectus muscle, the `up' map activates the
superior rectus. Note that `down' is not orthogonal either to `left'
or `right' because a saccade down and left is achieved by
simultaneously activating both the lateral and inferior rectus
muscles. However, the `up' map \e{is} orthogonal to the `down' map and
spans the edges of the surface where $\phi$ rolls over from 1 to
50. The line of maximum connection strength for the `up' map is along
$\phi=\mathrm{1}$; for `down' $\phi=\mathrm{26}$. Based on the training described in
Sect.~\ref{sec:methods:integrating}, the maps driving the superior
oblique (`Z+') and inferior oblique (`Z-') muscles were set to
1/10$^{\mathrm{th}}$ of the `down' and `up' maps.}
% From map_modelling2.m:
%    yh=0.0016*exp(0.067.*x);
%    yv=0.00195*exp(0.075.*x);

\subsection{Saccade accuracy} \label{sec:results:accuracy}

In Fig.~\ref{fig:saccvstarg}, we showed the result of running the model
to targets located on the principle axes, on which the model was
trained. We then simulated single saccades to targets in one hemifield
of the eye's field of view, with eccentricities between 6\dg~and
14.5\dg. As in the training, we ran the simulation 6 times for each
target, $\mathbf{\theta}^t = (\theta_{x}^t, \theta_{y}^t, 0)$ to
obtain mean saccade end-points.  Fig.~\ref{fig:errorsurfaceTM3} shows
saccade accuracy results for an entire hemifield in the na\"ive model
which passed the output of SC\_deep directly to SBG via the weight
maps. The ratio of the magnitude of the error vector to the magnitude
of the target vector is plotted using a colour map. This ratio is
shown for the full, three dimensional error vector in
Fig.~\ref{fig:errorsurfaceTM3}(a) and for the $x$, $y$ and $z$ components
in Figs.~\ref{fig:errorsurfaceTM3}(b)--(c).  Inspection of
Fig.~\ref{fig:errorsurfaceTM3}(a) shows that the end-point error is
minimal along the principle axes ($\theta_{x}^t=\;$0 or
$\theta_{y}^t=\;$0) and maximal near the 45\dg~oblique targets (blue
lines) with the end point error as high as 80\% of the programmed
saccade magnitude.  The $x$ component error map in
Fig.~\ref{fig:errorsurfaceTM3}(b) shows the same trend, mirrored about the
`Target X' axis, whereas the $y$ and $z$ component errors are,
relatively, much smaller.  Because the $x$ component of the error is
clearly contributing to end point errors which would not be considered
`on target', especially for oblique saccades, we considered the effect
of the non-uniform size of the hill of activity in SC\_deep.

In our model, the location, \e{size} and shape of activity in FEF, the
basal ganglia, thalamus and superior colliculus is eccentricity
dependent, in line with the retinotopic mapping stated by
\cite{ottes_visuomotor_1986}. More eccentric targets generate reduced
activity, because fewer retinal neurons are excited far from the
fovea. \cite{cope_basal_2017} showed that this relationship can
explain increased saccadic latencies for distal targets, resulting
from reduced activity in the decision making circuitry of the basal
ganglia.  However, the notion that activity in superior colliculus is
eccentricity-dependent conflicts with the result of
\cite{tabareau_geometry_2007}, who showed
that an invariant hill of activity was required if this complex
logarithmic weight mapping was to be used to drive a
two-degree-of-freedom saccadic burst generator, and also with
experimental findings, which do not show significant eccentricity
dependence, at least in the burst
layer~\citep{anderson_two-dimensional_1998}.
% Nice ref that - see Fig 16

To bring our model in line with these results, whilst maintaining the
eccentricity dependent activity in basal ganglia, we hypothesised that
a `widening projection' exists between two maps in superior
colliculus. \rvw{As described in
Sect.~\ref{sec:methods:integrating}, there is
now experimental evidence for similar projections
\citep{ghitani_excitatory_2014,bayguinov_hard-wired_2015} making this
a plausible suggestion.}  Activities in one SC\_deep layer remains
eccentricity-dependent, with loops back to thalamus and cortex and
through basal ganglia.  This activity is then fed through a
projection, which applies a Gaussian projection field, whose width
increases with increasing stimulus eccentricity according to
Eq.~\ref{eq:widening}.  The activity in this second SC\_deep layer is
then fed to the weight maps of the SBG. This model was called
`TModel4'. TModel4 was parameterised such that its horizontal and
vertical error was similar---so that its equivalent of
Fig.~\ref{fig:saccvstarg} showed a similar sum of squares error.

Figs.~\ref{fig:errorsurface}(a)--(d) show the \rvw{percentage errors for
TModel4}.  First of
all, note that the error magnitudes are much smaller. The mean errors
are smaller for every axis. The largest errors produced by the model
are approximately 15\%, which are within the boundaries of what some
authors regard as an accurate
saccade \citep{mcpeek_saccade_2002,mcpeek_incomplete_2006}.  The
magnitude of the largest error vector is approximately 1.5\dg.

This result indicates that the exponential part of the Ottes et
al.~weight map from SC to the SBG cannot on its own compensate for the
eccentricity-dependent size of the hill of activity. The introduction
of a widening projection field substantially improves the mean
accuracy of saccades across the field of view.  We therefore suggest
that the transformation between retinotopically mapped activity, and
eccentricity-independent activity width occurs within the superior
colliculus and works alongside a simple, monotonically increasing
weight map between SC and the SBG channels.

\subsection{Single saccades} \label{sec:results:singlesaccades}

Having finalised the model by setting the weight maps, we then
proceeded to exercise the model (TModel4), starting with saccades to a
single target; prosaccades. Fig.~\ref{fig:outmany}(a) shows 9
representative saccades to a single target luminance. Initially, the
eye had rotational state $\theta_x=\theta_y=\theta_z=0$ with its fovea
directed at a fixation luminance cross (span 6\dg, bar width 2\dg) of
magnitude 0.2 (in arbitrary units). At a simulation time of 0.4~s, the
fixation luminance was set to 0 and a target luminance cross of the
same dimensions as the fixation but with magnitude 0.3 was illuminated
at one of the 9 different locations, marked by crosses in
Fig.~\ref{fig:outmany}(a).  The resulting trajectories are plotted, with
colour indicating the relationship between trajectories and target
crosses. The approximate end-point error is visible in this figure,
although the last point in each trajectory is the saccade position at
0.8~s and not the velocity-based end-point described
above. Figs.~\ref{fig:outmany}(b) and (c) show the rotational components
of the blue and red trajectories in Fig.~\ref{fig:outmany}(a) along with
the target and fixation luminance values. Rotations are the eye's
Euler rotational components in the world frame of reference.

% These errors are larger than those in Thurat due to the hill of
% activity not being `perfect' Discussion point.

\subsection{Saccade Latencies} \label{sec:results:latencies}

To verify that our implementation of the brain model has the same
functionality as that reported in \cite{cope_basal_2017}, we
investigated the effect on saccadic response times of: target
eccentricity; and any gap or overlap between fixation off-time and
target on-time. We showed that the full model reproduces the `hockey
stick' shape shown in Fig. 7 of \cite{cope_basal_2017} and discovered in
experimental data \citep{reulen_latency_1984} for horizontal
[Fig.~\ref{fig:latvsall}(a)], vertical [Fig.~\ref{fig:latvsall}(b)] and
oblique saccades (not shown). The latency increases with eccentricity
far from the fovea because the retinotopic mapping reduces the
activity in the basal ganglia for more eccentric targets (see
Fig.~\ref{fig:mapping}). Closer to the fovea, the \rvwt{effect of} the
foveal mask \rvwt{on} the activity in FEF again leads to reduced input
into the basal ganglia and an increased time to achieve disinhibition
in SNr.

Fig.~\ref{fig:latvsall}(c) shows latencies achieved when varying the
time between fixation offset and target onset. This is termed
the \e{gap condition}; and is represented by a scalar value which, if
positive, refers to a gap between fixation offset and target onset,
and when negative, signifies an overlap, with the fixation luminance
persisting past the time at which the target is illuminated. A
negative gap is also termed an \e{overlap}.  Again, we verify the
behaviour presented in \cite{cope_basal_2017}, explained as resulting
from the inhibition of the cortico-thalamic loop by SNr. In the gap
condition, when the fixation luminance is removed, activity in STN
immediately begins to decay, allowing SNr activity to reduce and
thereby reducing inhibition on thalamus, allowing the target luminance
to build up quickly in FEF, thalamus and through the basal ganglia's
striatum and SNr. The shape of the curves in Fig.~\ref{fig:latvsall}(c)
matches the results in \cite{cope_basal_2017} for target luminances of
1 and 0.6; for overlaps longer than 100~ms (gap $<$ -100~ms), the
latency becomes constant; the saccade is programmed whilst the
fixation is present, with the target luminance inducing sufficient
activity in striatum to `break through' the SNr inhibition caused by
the fixation. If the target luminance is reduced to 0.3, the balance
is altered in favour of the fixation and the latency vs.~gap becomes
approximately linear and equal to the overlap time plus around 100~ms.

Fig.~\ref{fig:latvsall}(d) shows the effect of the dopamine parameter on
saccade latencies in gap, step and overlap conditions. In general, the
effect of decreasing the dopamine parameter was a smooth, monotonic
and undramatic increase in saccade latency. However, the data for the
overlap condition with a target luminance which was 3 times as bright
as the fixation luminance was more interesting. Here we see a
transition around a dopamine value if 0.7.  Below this value, the
basal ganglia is not able to select the target luminance until the
fixation is removed, reducing the excitatory drive from STN to SNr,
and consequently the inhibition from SNr to the thalamo-cortical
loop. For the target luminance 0.6, 0.7 dopamine allows the basal
ganglia to select sufficiently well so that the target can build up in
the thalamo-cortical loop, in spite of the fixation
overlap.

The relationship between latency and the target luminance is given in
Fig.~\ref{fig:latvsall}(e). This shows latency for a 100~ms gap, step
and 100~ms overlap conditions for a given fixation luminance of 0.2,
and a horizontally located target at $\theta_{y}^{t}=$-10\dg. For the
gap condition, we see very short latencies for luminances of about
0.75 and above.  Finally, the activity driving these express saccades
is initiated by high firing rates in the superficial layer of SC
(SCs), which then drives activity in thalamus and through the basal
ganglia. A gradual transition from express saccades to reflexive
saccades is observed as the contribution of the SCs becomes weaker and
the drive from FEF into the thalamo-cortical loop becomes necessary to
elicit a saccade.
%
A similar gradual transition, albeit for higher latencies is seen for
the step condition. At higher target luminances, the SCs has a greater
effect on the activity in the thalamo-cortical loop. However, the
activity in STN caused by the fixation luminance increases the latency
at all luminance values compared with the gap condition.
%
The overlap condition leads to increased latencies for luminances
below 2.5, but meets the step condition above this value, at which the
0.2 fixation luminance appears to have a negligible effect on the
system.

\subsection{Saccade sequences} \label{sec:results:sequences}

We now present results derived from the fully parameterised and
integrated model; where we took advantage of the fact that it is a
closed loop system. This allowed us to present sequences of target
luminances and allow the model to direct its fovea at the most salient
target.

\subsubsection{Out \& return}

We investigated the behaviour of the model for saccade sequences.  In
one experiment, we illuminated a fixation cross from 0~s until 0.4~s,
followed by a target at (0,-10\dg) from 0.4~s until 0.8~s. Finally,
the fixation was again shown from 0.8~s until the end of the
simulation at 2~s. This induced a saccade to a 10\dg~eccentricity,
followed by a return saccade back to the null point. We noticed some
irregularities in the return saccades, which \rvw{were
accurate}, but had a significant overshoot.  More perplexingly, if the
target was switched repeatedly between 0\dg~and 10\dg, second and
subsequent \e{outward} saccades also showed this overshoot. We found
that the cause of these irregularties was the lack (in `TModel4') of
any mechanism to reset the tonic neurons in the SBG after the first
saccade. This resulted in TN activity in the left channel \e{and also}
in the right channel. Interestingly, this ensured that, at least for a
few, consecutive out-and-return saccades, the saccade accuracy was
accidently relatively good, with trajectories resembling experimental data
(\cite{bahill_trajectories_1979}, p.~6). \rvw{Had the return saccades not
been so accurate, we may have noticed the lack of a tonic neuron reset
mechanism and corrected this oversight earlier.} Such a mechanism
is indeed proposed and included in the connectivity of the
\cite{gancarz_neural_1998} model. We implemented this feature by adding an
additional inhibitory input to the `integrator' SBG component of TModel4,
driven by the contralateral EBN population, naming the new model
`TModel5'.  Now, when the eye is directed towards an eccentric target
which is then exchanged with a target at the null point, the EBN
activity toward the null point will tend to extinguish the TN activity
which was holding the eye at the eccentric position.  We verified that
none of the single saccade results were affected by this modification.
% This doesn't account for *all* cases where you'd want to remove the
% TN activity. Imagine a saccade purely UP then a saccade to a
% location purely RIGHT. You'd need to remove the TN activity on the
% UP channel, but you'd not be inducing any DOWN with the new target.

Fig.~\ref{fig:outrtn} shows the outward and return trajectories produced
by the experiment with the TN reset mechanism. Panel (a) shows the $x$
and $y$ rotation trajectory; panel (b) shows individual rotational
components of the eye. Fig.~\ref{fig:outrtn}(c) shows out and return
trajectories for three other saccade targets; horizontal, vertical and
oblique. The trajectories have characteristic shapes and also show
some stochastic variation caused by the noise in the model [see dashed
trajectories in Fig.~\ref{fig:outrtn}(a)].

The return trajectories (magenta lines) showed a distinctly different
form from the outward trajectories. They overshot their destination
(the null point) significantly. This resulted from the removal of the
TN activity which was holding the eye at the eccentric target
location. Removal of this activity, and thus the static force exerted
by the corresponding extraocular muscle, meant that the eye was
subject both to a new muscular force towards the null point
\e{alongside} the restorative spring force of the lengthened rectus
muscle. This stands as a shortcoming of the model.

\subsubsection{Double steps}

In another experiment, we probed the response of the model to double
step stimuli of the type described in \cite{becker_analysis_1979}. In
that work, the response of human subjects was investigated when shown
stimuli at 15\dg~and 30\dg~eccentricity with variable delay between
the stimuli. If the smaller eccentricity stimulus was shown first,
followed by the more distal on the same side of the field of view,
this was called a `staircase' presentation.
% If the more distal stimulus preceded the 15\dg~stimulus (again with
% both in the same hemifield), this was called `pulse undershoot'.
We carried out a `staircase' presentation, shown in
Fig.~\ref{fig:doublesteps}, where our small eccentricity luminance was
at 8\dg~and our more distal luminance was at 12\dg~(both to the right
of centre). \rvwt{The stimuli could not be presented at 15\dg~and
30\dg~to match the experiment, because 30\dg~saccades were outside the
range of the model.}

We found that there was a critical time delay between the
luminances of about 30~ms. If they were presented with a delay smaller
than this value, then a single, slightly hypermetric saccade was
made. This response type is called a \e{final angle response}. A delay
greater than 30~ms between the stimuli would lead to double step
saccades (a so-called \e{initial angle response}), with the first
saccade arriving at 8\dg~(though with greater variability than
normal), and a second saccade being made to a location hypometric of
12\dg~after a pause of about 240~ms.  Fig.~\ref{fig:doublesteps}(a) shows
the mean trajectories from 5 simulations of the staircase doublestep
presentation alongside the result for a single saccade to the final
angle of 12\dg. Dash-dot lines show $\pm$1 standard deviation about
the mean. The corresponding trajectories are shown in
Fig.~\ref{fig:doublesteps}(b).

Inspection of the activity maps in FEF and SC\_deep (not shown)
indicates that when the 8\dg~target is illuminated for 30~ms or more,
the activity associated with this target angle is able to dominate the
activity, hence the execution of a reasonably accurate saccade. The
inhibitory feedback from the SBG then extinguishes activity in FEF,
thalamus and SC, which means that a full 200~ms or more is required to
allow activity in these populations to build up again in order to make
the smaller saccade from 8\dg~to 12\dg. This is in contrast to
experimental findings in which the corrective second saccade is often
executed \e{more quickly} than if it were programmed on its
own~\citep{becker_analysis_1979}.

\section{Discussion} \label{sec:discussion}

The aim of this study was to demonstrate the importance of modelling
neurological systems \e{in concert with} the biomechanical systems
with which they have evolved. We hypothesised that by
combining existing neurophysiological models with an accurate model of
a musculo-skeletal system, and then \rvw{closing the
`agent-environment-agent' loop} by allowing the
movements of the virtual muscles to modulate sensory feedback to the
brain model, shortcomings in the constituent models would be revealed,
leading to new knowledge.

To demonstrate the validity of this closed-loop
approach, we built an integrated model and then identified the
modifications which were necessary to give it the ability to make
accurate movements under one type of stimulus. We then examined its
behaviour with other stimuli. We chose the oculomotor model as a basis
for this study because it has only three degrees of freedom, making it
one of the simplest musculo-skeletal systems. Furthermore, eye
movements fall into several well-defined categories, each being
controlled by separate brain circuits, we were therefore justified in
modelling a system which produced only saccadic eye
movements. Nevertheless, we are aware that we did not create a
complete model of the system; no treatment of the cerebellum was
attempted, justified because cerebellum appears to have only a minor
effect on saccade accuracy \citep{dean_adaptive_2008}, probably
correcting for slow to medium timescale changes in the physical
dynamics of the eyeball \citep{dean_learning_1994}.

To summarise our model integration: We combined
the \ccg~model \citep{cope_basal_2017} with a saccadic burst generator
model based on the work of \cite{gancarz_neural_1998}, using this to
drive the input of a new biomechanical eye model. To achieve the
spatial transformation from the retinotopic maps of the \ccg~model to
the six `muscle channel' inputs for the saccadic burst generator, we
used the mapping of \cite{ottes_visuomotor_1986} to produce
parameterised weight maps along with an empirically discovered synergy
for the torsional weight maps. We introduced an additional
transformation to the brain model to achieve invariant sized hills of
activity in superior colliculus to fulfil the invariant integral
hypothesis of \cite{tabareau_geometry_2007}. We closed the loop using
a software component which transformed a view of a world containing
luminous cross shapes into the eye's frame of reference, given its
instantaneous rotational state. This component also computed the
inverse of the mapping from \cite{ottes_visuomotor_1986} to project
the view retinotopically into the brain model.
%
This paper serves to describe how we achieved the integration in order
to test our hypothesis, and we intend that the material and methods
section, along with the model code itself, will help others to carry
out similar studies. We will discuss what can be learned from an
integrated model of a combined brain and biomechanical system, using
our oculomotor system as an example and then consider how this study
compares with other modelling and robotic studies of closed-loop
systems.

Our integration approach revealed three ways in which this model fails
to provide a full understanding of the saccadic system. In each case,
the issue is made clear \e{as a result of the integration}. This is
not to say that other approaches may not also reveal shortcomings; we
will see that one of our cases has been independently identified
\citep{groh_effects_2011}.

\subsection{The need for a widening projection field}

The original combination of the \ccg~model with the theoretical weight
maps of \cite{ottes_visuomotor_1986} and \cite{tabareau_geometry_2007}
resulted in a model which was able to produce accurate saccades only
along the principle rotational axes
(Fig.~\ref{fig:errorsurfaceTM3}). Thus, \e{the integration of the models}
suggested that an additional layer was required to achieve accurate
saccades for oblique, as well as for horizontal and vertical
saccades. Although the \e{need} for an invariant integral is discussed
in \cite{tabareau_geometry_2007} as resulting from their theoretical
study, the mechanism by which such an invariant Gaussian hill is
generated is not. By combining the models, we were forced to consider
this mechanism, and hypothesised that a widening projection field
would be a candidate mechanism. The results of Fig.~\ref{fig:errorsurface}
indicate that a substantial improvement in accuracy is indeed achieved
by this new mechanism.

\subsection{Saccades from non-null starting positions}

The implementation of a biologically accurate model of the eye, and
the closed-loop nature of the model makes it very natural to consider
how the model will behave when making saccades from arbitrary starting
positions, or how it would respond to a sequence of stimuli. This was
the motivation for the out-and-return experiment (Fig.~\ref{fig:outrtn})
as well as for the double step experiment (Fig.~\ref{fig:doublesteps}).
We found that return saccades were substantially affected by the
biomechanics of the eye, as the brain and brainstem model had no
mechanism to account for the position-dependent restoring forces
applied by the eye. This question has been addressed by other authors;
\cite{groh_effects_2011} investigates the effect of
initial eye position on stimulated saccades and finds a need for the
signal in superior colliculus to be modulated by an eye position
signal. \cite{ling_effects_2007} shows the existence of a position
dependent firing rate offset in abducens neurons. Though we will not
speculate here on the mechanism by which return saccades may be made
accurate whilst also resetting the activity of tonic neurons in the
SBG, it is interesting that in the model in which we omitted to reset
TN activity (TModel4), we obtained relatively accurate out-and-return
saccades which closely resembled experimental data.  We suggest that
residual activity in TN populations may offer an explanation for how
the restorative force exerted by the elastic oculomotor muscles is
compensated for. A comparison of this idea with that
of \cite{groh_effects_2011} (that there is a modulation, from a
brainstem signal, of the SC readout) would make a subject for a future
study. Although these existing studies have highlighted this issue,
the inaccurate return saccades which the model makes from eccentric
starting positions provide a clear example of the way in which
integrating known models into a closed-loop system can highlight
deficiencies in the model.

\subsection{Inhibitory feedback from saccadic burst generator to brain}

The third issue raised by the integration of the component models of
the saccadic system has, like the return saccades, to do with
resetting activity. In this case, rather than the reset of activity in
the TN population in the brainstem, it is the question of how the
activity in the \e{brain} model should be reset after each
saccade. When a target luminance is projected onto the World
population in the model, this induces activity which `reverberates' in
loops through FEF, basal ganglia, SC and thalamus. The brainstem
contains a mechanism to limit the timescale of a saccade (inhibitory
feedback from EBN, via IBN to LLBN; see Fig.~\ref{fig:sbg}). However, if
the activity in SC is not reset, then following the completion of the
first saccade, a series of subsequent `staircase' saccades will be
executed. There needs to be a mechanism to extinguish activity in SC,
but also in FEF and thalamus, as activity in either of these
populations can build up and eventually cause repeat activity in SC
and another saccade. We added hypothetical inhibitory feedback
connections to our model, such that the IBN populations in the SBG
would inhibit activity in FEF, thalamus and SC\_deep (Fig.~\ref{fig:sbg}),
preventing the occurrence of staircase saccades.

An examination of the behaviour of the model when presented with
`double-step stimuli' reveals a problem with our inhibitory feedback
connections. We found that when double-step stimuli were presented (where
an initial target at 8\dg~was replaced with a 12\dg~target after 30 or
40~ms) and a double saccade was made [Fig.~\ref{fig:doublesteps}(a), black
lines] the second saccade latency was \e{longer} even than the initial
saccade. This contrasts with \cite{becker_analysis_1979} who find that
second, corrective saccades occur with \e{shorter} latencies.  This
suggests that the inhibitory reset signal implemented in this model is
too strong or has the wrong timescales. This issue highlights the fact
that connections \e{between} component models are quite as important
as the connections within each model.

% Inhibition experimental evidence
\rvw{There is some evidence for an inhibitory projection to SC from the
brainstem. \cite{corvisier_possible_1991} offer evidence for a
projection from the propositus hypoglossi nucleus. This lies upstream
from motoneurons and (in primates) encodes eye
velocity \citep{dale_nucleus_2013}, rather than head movement
velocity. Although the propositus hypoglossi does not lie in exactly
the same functional location as our IBN population (instead it sits
between TN and MN), it offers a possible inhibitory feedback signal
proportional to eye velocity and may help to reduce activity in SC
post-saccade.}
%
\rvwt{Alternatively, it is possible that activity in
FEF and thalamus are reset via a `timed signal'. Feasibly, after
activity in FEF exceeds a threshold, an internal, inhibitory feedback
signal could be activated. This inhibition should have a timescale of
sufficient duration to reduce activity in FEF, thalamus and, via an
increase in inhibitory output from SNr, also in SC. Indeed, the
cortical microcircuit contains a variety of morphologically distinct
GABAergic neurons \citep{douglas_neuronal_2004} which could fulfil
this functionality. A similar mechanism would then be required in SC,
to reset activity generated by direct excitation via the
retinal-collicular pathway which generates express saccades. Again, SC
is a multi-layered structure, containing GABAergic interneurons
\citep{munoz_lateral_1998,meredith_intrinsic_1998,helms_organization_2004,sooksawate_organization_2011}
and there is mounting evidence that saccade dynamics are generated
within
SC \citep{kaneda_regulation_2008,goossens_optimal_2012,bayguinov_hard-wired_2015}. Thus,
a more complex treatment of the SC and FEF regions in the model may
well obviate the need for inhibitory feedback from brainstem to SC,
FEF and thalamus.}

%
% A couple of last points to round up the discussion of THIS model
%

% Reviewer 1's point 2.
\rvw{Considering whether a feedback connection, or internal,
recurrent inhibition is responsible for activity-reset in the brain
model raises a more general question about modelling the central
nervous system. We should consider whether inaccuracies within one
part of the model may propagate errors through the closed-loop system
that cannot be counteracted by another part of the simulation. There
is no way to know, from integrating sub-systems, which properties hold
true, and which are false. However, by integrating models and
examining the behaviour of the combined model, we are presented with
the right questions to ask of the model and the experimental data. In
the case of activity-reset, this is to re-assess whether there exists
inhibitory feedback from brainstem to the SC and FEF regions, and to
find out how an integrated model with self-regulatory mechanisms in SC
and FEF may perform.}

% In a model of SC, \cite{massone_neural-network_1994} suggests that
% internal mechanisms, learned via a supervised method, are sufficient
% to generate dynamics.

% Cerebellum omitted
The omission of the cerebellum will not have escaped the reader's
notice.  Whilst many of the nuclei known to be involved in the
production of saccadic eye movements are incorporated within the
model, the cerebellum is not. The cerebellum is known to play an
important r\^ole in saccade programming
\citep{dean_learning_1994, schweighofer_model_1996,quaia_extent_2000,kleine_saccade-related_2003}.
It may be able to completely replace the functionality of the
colliculus when
lesioned \citep{aizawa_reversible_1998,lefevre_distributed_1998}.
However, this r\^ole is typically considered to be one of accuracy
tuning
\citep{barash_saccadic_1999,dean_learning_1994}; operating as an additive
model. Furthermore, saccades made by individuals with cerebellar
ataxias perform with only moderate loss of saccade accuracy
\citep{barash_saccadic_1999,federighi_differences_2011}. Because we
did not address learning in our model, and because our aim was to
demonstrate the utility of integrating brain with biomechanics in
order to highlight deficiencies, we considered the omission of the
cerebellar nuclei acceptable in the present work.

% Component stretching/main sequence not addressed
We have not addressed the question of saccade duration in this paper.
Saccade duration is of interest in models which produce two (or three)
dimensional saccades, because the dynamics of a saccade follow well
known relationships with the saccade eccentricity, regardless of the
saccade angle. This causes a problem for models (such as the present
one) for which some of the dynamic behaviour is generated within
orthogonal components.  For example, saccade duration increases with
target eccentricity. A 10\dg~eccentricity oblique (45\dg~up and right)
saccade is composed (approximately) of a 7\dg~upwards component and a
7\dg~rightwards component. If the component based model is responsible
for the dynamics, then the 10\dg~oblique saccade would be expected to
have the dynamics of a 7\dg~up or 7\dg~right saccade. This is not
found in practice, and the components are said to have been stretched,
hence the name for this effect `component
stretching'. The \cite{gancarz_neural_1998} model is reported to take
account of the component stretching effect via the OPN neuron
population. We did not find this effect in our implementation of the
model; the duration of oblique saccades at a given eccentricity was
always substantially different from the duration of the corresponding
purely vertical or horizontal saccade. Because there is a somewhat
complicated interplay between the dynamics of the superior colliculus
driving the dynamic system of the SBG, we feel this is outside the
scope of the current work and a subject for a future paper.

\subsection{Comparison with other studies}

\rvw{We have called this closed-loop, biomimetic modelling approach \e{computational
neurobehaviour}, in which a complete, behaving model is constructed,
with attention paid to the biological accuracy of each brain and
biomechanical sub-system. There already exist many
closed-loop \e{robotic} systems which receive sensory input from the
world, process that input and generate behaviour by activating motor
systems
\citep{yu_development_2004,fend_artificial_2004,pearson_whiskerbot:_2007}. We
now consider whether robotic systems which model biological
components \e{in hardware} could fall within our new category.
%
Using a number of examples, we will attempt to illustrate what we mean
by computational neurobehaviour. We'll consider which examples fall
into the new category and which are covered by other fields of
robotics or computational neuroscience.}

\rvw{\cite{pearson_whiskerbot:_2007} describes a wheeled robot which has a
biomimetic whisker sensory system, along with a biomimetic neural
system imitating the operation of the rat's sensory processing and
controlling the movement of the robot. \cite{fend_artificial_2004} is
a similar, wheeled, whiskered robot, with a repertoire of three
behaviours organised in a subsumption architecture. In both robots,
actions that are selected within the brain model drive a
non-biologically accurate wheel motor control algorithm to achieve
rotational and translational movements. Although both have sensory and
processing systems which are guided by biology, the non-biological
motor control stage prevents us from considering these as being
studies of computational neurobehaviour. Instead, we would refer to
these as \e{embodied models}, as described
in \cite{bolado-gomez_biologically_2013}, a study in which the
learning behaviour of a biomimetic `core' model is embedded within an
engineered `architecture' (a wheeled robot) which closes the
agent-environment loop.}
%
\rvw{\cite{yu_development_2004} report on a biomimetic fish robot,
whose motor system closely resembles that of the real fish. The robot
is able to operate in a closed-loop mode, where sensory input is
provided to the non-biomimetic control algorithms from overhead
cameras, but its control system is also able to operate in open-loop
mode. Neither the control system, nor the sensory system are
biologically accurate and we would not describe this study as
computational neurobehaviour. Nevertheless, the biomechanically accurate
motor system they describe has the potential to form part of a
computational neurobehavioural study of swimming behaviour in the
fish, if it were combined in a loop with suitable sensory input and
sensory processing models.}
%
\rvw{\cite{knips_neural_2017} is a report of a reach-and-grasp robot
arm controlled via a dynamic neural field brain model. The sensory
input for this system---its `eyes'---is a Microsoft Kinect sensor; it
also has somatosensory feedback from the fingers of the robot's
hand. The neural field `brain' controls the seven degrees of freedom
of the arm to carry out the reach-and-grasp action. While this robot
has closed-loop control and is clearly inspired by biology, it remains
a study of robotics and of the improvement of the control of the
robot's reach-and-grasp function, rather than a study which aims to
learn more about the biology of a primate arm. For this reason, we
would describe the study of \cite{knips_neural_2017} as an embodied
model.}

\rvw{To summarise, in most closed-loop robotic studies which
incorporate neuromimetic models, the hardware forms an `engineered
surround architecture', allowing for the examination of the behaviour
of the embodied model. However, suitably \e{biomimetic} hardware such as
the fish in \cite{yu_development_2004} would not be excluded from
computational neurobehavioural studies.}

\rvw{Modern programming platforms, often originating from the
computer game industry, make it relatively easy to model a virtual
environment. Consequently, an increasing number of studies into
robotic or neuromimetic control are carried out with virtual robots
operating within a virtual environment.}
%
\rvw{\cite{nguyen_saccade_2014} and
\cite{thurat_biomimetic_2015} are two studies of the oculomotor system
which model sensory input, neural control \e{and} motor output in
software. These studies fall outside the remit of computational
neurobehaviour only because they omit to close the sensory
loop. \cite{dewolf_spiking_2016} describes a reach model comprising a
simplified virtual arm (with fewer degrees of freedom than a primate
arm), and a biologically inspired brain model. This model also omits
to close the sensory feedback loop and we consider it a computational
neuroscience study of a (virtually) embodied model.}
%

% Reviewer 1's point 4.
\rvw{There are also experimental closed-loop approaches to
understanding sensorimotor control. \cite{ejaz_closed-loop_2013}
places a fly in a fixed position, and couples it with a free-to-move
robot. The sensory input collected by the robot is projected onto the
eyes of the fly, and activity from a selected neuron in the fly's
brain is used to drive a control system for the robot's
movements. This allows the experimenters to study the behaviour of the
fly's brain operating in a closed-loop condition that is more natural
than the open-loop condition that many other experimental techniques
mandate. The results from closed-loop experiments will undoubtedly
inform future neurobehavioural models.}

%
% Now the closing section of the discussion.
%

\rvw{Thus, while there are many models that close the
agent-environment loop and display partial biological plausibility,
the biomimetic features are usually confined to a sub-system of the
entire model. This leads us to formalise a definition of computational
neurobehaviour as: \e{The study of biological sensory-motor system
behaviour using biologically accurate models of sensory input, brain
and motor sub-systems operating in a closed-loop}.}
%
% Another computational neurobehavioural study?
\rvw{We believe our oculomotor model is one of the first such models
using this approach and shares many features with that
of \cite{arena_motor-skill_2017} which describes a robotic insect
system based on the fly species \e{Drosophila Melanogaster}. It has
biomimetic insect legs implemented in a virtual robot and a
neuromimetic brain. Closing the loop is a visual sensory input system
which is able to determine the distance between the robot and an
obstacle and the obstacle's height. The authors demonstrate that the
virtual robot is able to learn to climb obstacles in a realistic
manner and suggest that the model may be compared with experimental
data from future Drosophila experiments addressing obstacle climbing
and learning. The visual system is not described in detail, but if it
is modelled in a biologically plausible manner, then this work may
reasonably be described as a computational neurobehaviour study,
contemporaneous with our own.}

% Significance of this work.
%% This work represents a step forward in the modelling of neuromuscular
%% systems, not because it significantly advances any of the constituent
%% models, but because it \e{integrates} the models into a
%% complete, \e{behaving} system.  This is not the first integrated brain
%% model composed of separately developed components. The works
%% of \cite{nguyen_saccade_2014} and
%% \cite{thurat_biomimetic_2015} are both based
%% on an example of a brain model which drives a simple, second order
%% model of the eye. \cite{dewolf_spiking_2016} describes a reach model
%% which integrates models of cortex and cerebellum to drive a two
%% degree-of-freedom arm model. Both of these example systems
%% nevertheless operate using `curated' inputs supplied by the modeller.

%% In contrast, the current work allows the state of the system to
%% determine the input delivered to the model. The modeller only curates
%% the state of the world at each time point, but the actual input to the
%% model depends on the eye's rotational state.  This is, to our
%% knowledge, the first model which integrates the brain with an accurate
%% biological system and closes the loop in this way, enabling the
%% system to reproduce behaviour.  As such, it offers a platform for
%% testing more complex saccadic behaviour such as antisaccades or
%% saccades in the presence of distractor stimuli. We believe that by
%% building closed loop systems which express behaviour, we, and others
%% will develop a new field of \e{computational neurobehaviour}, which
%% will share themes from neuroscience, artificial intelligence, decision
%% science and embodied robotics.

\section*{Disclosure/Conflict-of-Interest Statement}

The authors declare that the research was conducted in the absence of
any commercial or financial relationships that could be construed as a
potential conflict of interest.

\section*{Author Contributions}

SJ, AB and AC implemented existing parts of the model in SpineML. AB
developed the saccade generator brainstem model. SJ performed the
technical and scientific integration of the biomechanical eye.  CP and
KM developed the biomechanical eye model. SJ wrote the manuscript; SA,
AB, KG and KM contributed to the manuscript. KG conceived the project.

\section*{Acknowledgements}

\textit{Funding\textcolon} European Union FP7 Grant 610391 `NoTremor'.

\section*{Supplemental Data}

The model specification, results and all code required to reproduce
the results of this work are available at:
https://github.com/ABRG-Models/OMM\_NeuroMuscular

\selectlanguage{English}
\bibliographystyle{frontiersinSCNS_ENG_HUMS}
% The bibliography NoTremor.bib is the one exported from Zotero.
\bibliography{NoTremor}

\section*{Figures}

% Figure 1
\begin{figure}[h!]
\begin{center}
\includegraphics[width=\textwidth]{./figures/ccg_model.png}
\end{center}
\caption{ \bol{a) The macroscopic architecture of the \ccg~model.} The main
nuclei modelled as brain systems are: basal ganglia (BG), frontal eye
fields (FEF), thalamus (THAL) and superior colliculus (SC). The
retinal input is presented via non-biologically plausible retinal
populations. The loops through basal ganglia, which define the
architecture, are shown with coloured lines: the cortical loop
(through FEF and THAL) in green and the sub-cortical loop (through SC
and THAL) in blue. Connections with arrowheads indicate excitatory
connections, those with circles are inhibitory. A red border indicates
that the box represents a sub-system of two or more populations; a
black box indicates (at least, within the context of the model) a
single neural population. The BG box is expanded in: \bol{b) The basal
ganglia model component.}
%
This shows a basal ganglia comprising striatum (Str\_D1 \& Str\_D2),
subthalamic nucleus (STN), globus pallidum externum (GPe) and
substantia nigra reticulata (SNr). The model has three action channels
shown as black boxes within each blue population border. Three
channels of cortical input to the BG are also depicted. Red indicates
the activation level of a given channel, helping to illustrate the
selection mechanism. For example, the channel indicated by the
leftmost bar has a high salience (cortical input) and excites activity
in Str\_D1 which then inhibits the leftmost bar in SNr. The diffuse
projection from STN is equivalent to summing its projections
channel-wise, and then projecting the sum to all channels of its
target populations (the blue arrows indicate that all channels of GPe
and SNr are targeted by the connection). Dopaminergic modulation of
the inputs to the striatum are indicated by the blue circles labelled
`d' and the dotted lines. The SNr sends inhibitory output projections
to its targets. \bol{c) 2D Gaussian weights.} The `GaussianKernel'
connectivity pattern is based on the `in-plane displacement' between
the location of a neural element in one layer, and the location of a
target neural element in the target layer for the connection. The
potential weight of the connection is given by a 2D Gaussian function,
which is maximum for the target neuron which exactly corresponds to
the source neuron, and drops down for target neurons which are
horizontally displaced from the source neuron. A threshold is applied
to avoid a computationally expensive all-to-all connectivity (with
most of these connections having negligible weight values).  If the
weight is non-zero, then a connection is made from source to target,
otherwise no connection is made. \bol{d) Gaussian connectivity.} This
image shows connectivity (green rays) from two source neurons (in
Str\_D2, brown spheres) to target neurons (in GPe, blue spheres). The
circular connectivity pattern is seen. This does not show the weight
\e{values}, which reduce `towards the edge of the circle' and follow the
relationship shown in (c).}
\label{fig:ccg_model}
\end{figure}

% Figure 2
\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.9\textwidth]{./figures/Brain_Model.png}
\end{center}
\caption{ The brain model. This is the SpineCreator `network layer' view of
the model. Each box represents a neural population with 2500 elements,
arranged in a 50$\times$50 grid. The SpineML component name is printed
on the bottom right corner of each population box and the population
name is at the top. The overall connectivity between populations is
represented by the projection arrows with the colour indicating the
connectivity scheme (one-to-one connections are red, Gaussian kernel
connections are dark blue and so on). Excitatory connections have
arrowheads and inhibitory connections have circles, although for
details of the behaviour of the connections, the weight-update and
post-synapse components must be studied. Briefly, the model comprises
a \e{World} population, into which a retinotopically organised view of
the world is introduced. This information is passed into cortical
populations (FEF) and subcortical populations (SC) via a simple model
of the retina. These feed a cortico-thalamo-basal ganglia loop, which
selects which region of the deep layer of superior colliculus should
be disinhibited, allowing activity to build up therein. The five
populations comprising the basal ganglia are enclosed in a grey
outline. Note that substantia nigra pars compacta is not modelled
here, instead the level of dopamine in the striatum is set via a
parameter in the Str\_D1 and Str\_D2 populations}
\label{fig:brainmodel}
\end{figure}

% Figure 3
\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.9\textwidth]{./figures/mapping.png}
\end{center}
\caption{ Representative mapping from eye's frame of reference in Cartesian
co-ordinates to retinotopic co-ordinates. (a) The mapping of
luminances in the eye's frame of reference. The world input is
pre-defined by a JSON configuration file. Luminance position, size and
shape can be defined in this file, along with the times at which
luminances appear and disappear. The worldDataMaker.cpp code computes
the locations of the luminances in the eye's frame of reference, given
its rotational state. It also computes a 2D Gaussian convolution of
the luminances. Here, there are two cross shaped luminances spanning
10\dg, one of value 0.8 at the fixation point (0,0) and one of value
0.5 at a peripheral position (0,-12\dg). Note that these crosses have
the same `bar width' of 2\dg~as the crosses used in the simulations,
but their span of 10\dg~is greater than the 6\dg~used in the
simulations, to make these images clearer. (b) The locations of the
luminances in the eye's frame of reference are then converted into
retinotopic co-ordinates, with centrally located luminances being
represented at low values of $r$ and more peripheral luminances having
higher values of $r$. $\phi$ encodes rotational angle: 1 and 50 encode
upward movement; 13 is left; 25 is down; 37 is right. The output of
the World component is fed into FEF\_add\_noise and into the retinal
neuron populations. The colour map makes it possible to distinguish
between the two crosses. (c) The FEF\_add\_noise populations adds a
level of noise to the signal representing processing of the signal in
visual cortex. (d) A Gaussian projection from FEF\_add\_noise to FEF
further blurs the activity in FEF. FEF is the input to the basal
ganglia and one input to superior colliculus.}
\label{fig:mapping}
\end{figure}

% Figure 4
\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.6\textwidth]{./figures/Brain_Stem_1channel.png}
\end{center}
\caption{ One pair of channels of the saccadic burst generator (SBG) for left
(cyan) or right (green) movements. Collicular activity in SC\_avg
excites the channels via SBG weight maps. Each box represents a neural
population and shows the population name, the number of neural
elements (here 2500 or 1) and the SpineML component name; \e{LIN} for
Leaky integrator or \e{integrator}. Key: LLBN: Long lead burst
neurons; IBN: Inhibitory burst neurons; OPN: Omnipause neurons; EBN:
Excitatory burst neurons; TN: Tonic neurons; MN: Motoneurons.}
\label{fig:sbg}
\end{figure}

% Figure 5
\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.5\textwidth]{./figures/eye_screen_plus.png}
\end{center}
\caption{ The biomechanical eye. a) an OpenSim rendering of a pair of
biomechanical eyes showing the positions of the extraocular
muscles. Note that i) volume visualization of muscles should not be
confused with FEM muscle models; it is provided for user feedback
purposes, i.e. shape and color change depending on the muscle
activation, ii) superior and inferior oblique are visualized up to
their respective muscle pulleys. b) OpenSim rendering of biomechanical
eye within a head model c) Top-down schematic cross-sectional view of
the biomechanical eye situated within a spherical screen, with a
horizontal rotation towards a luminances at an angle of $-\theta_{y}$
about the $y$ axis. The $y$ axis points up, out of the page.}
\label{fig:bioeye}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.9\textwidth]{./figures/model_framework.png}
\end{center}
\caption{ The model framework. The model is specified using a combination of
declarative XML files and hand-coded C++. These original model
specifications are shown within the blue box. b) The green box shows
the processes which are applied to the model specification to produce
the BRAHMS system. Most of the process is defined within the scripts
which make up \stob, but the hand-written components must be manually
compiled and installed within the BRAHMS Namespace, allowing the
BRAHMS executable to locate them at runtime. c) The red box shows the
resulting BRAHMS system ready to be executed by the BRAHMS
executable. In practice, this call is made by \stob.}
\label{fig:modelframework}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.9\textwidth]{./figures/scdeep_plus.png}
\end{center}
\caption{ a) \& b) Tabareau-style weight maps for `left' and `down' components
of a saccade. c) The two weight maps in a) and b) shown on the same
graph, viewed from above. Circles show the locations of potential
hills of activity. Purple, dashed circles encode saccades down;
red circles encode saccades with both a left and a down component. d)
Showing the additional deep layer of superior colliculus (SC\_deep2)
and the output layer (SC\_avg, named for the fact that in an earlier
version of the model, it received the output of the centroid of
SC\_deep).  The widening Gaussian projection is shown as the arrow
between SC\_deep and SC\_deep2.}
\label{fig:scdeep}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.8\textwidth]{./figures/sacc_vs_targetpos.png}
\end{center}
\caption{ Accuracy at different target eccentricities for fixation luminance
0.2 and target luminance 0.3.}
\label{fig:saccvstarg}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=\textwidth]{./figures/weightmaps.png}
\end{center}
\caption{ Weight maps for the connections between the output layer of superior
colliculus and the six long lead burst neurons of the saccadic burst
generator model. Each map increases exponentially with increasing $r$,
multiplied by cosine($\phi$) about its `active' axis.}
\label{fig:weightmaps}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.95\textwidth]{./figures/errorsurface_TModel3.png}
\end{center}
\caption{ The end-point error surface for the original, na\"ive model
(TModel3). a) The ratio of the magnitudes of the total error vector
and the target vector, expressed as a percentage. b) The ratio of the
magnitude of the $x$ component of the error vector to the magnitude of
the target vector, expressed as a percentage. c) As (b) but for the
$y$ component. d) As (b), for $z$ component. All colour maps are shown
with the same scale. The target rotations, $\theta_{x}^t$ and
$\theta_{y}^t$ are denoted `Target X' and `Target Y' in the figure.}
\label{fig:errorsurfaceTM3}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.95\textwidth]{./figures/errorsurface.png}
\end{center}
\caption{ The end-point error surface for the model in which a widening
projection field was added to the model of the superior colliculus. a)
The ratio of the magnitudes of the total error vector and the target
vector, expressed as a percentage. b) The ratio of the magnitude of
the $x$ component of the error vector to the magnitude of the target
vector, expressed as a percentage. c) As (b) but for the $y$
component. d) As (b), for $z$ component. All colour maps are shown
with the same scale. The target rotations, $\theta_{x}^t$ and
$\theta_{y}^t$ are denoted `Target X' and `Target Y' in the figure.
Note that the range of the colour scale is 0 to 20\%, a much smaller
range than the range in Fig~\ref{fig:errorsurfaceTM3}.}
\label{fig:errorsurface}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.8\textwidth]{./figures/outmany.png}
\end{center}
\caption{ Representative single saccades. a) Trajectories from 9 saccades to a
single target at 9 different locations. In each case, a fixation cross
luminance of magnitude 0.2 was displayed at (0,0), the start position
of the eye, until time 0.4~s. The target luminance, magnitude 0.3 was
illuminated at time 0.4~s. Trajectory shape is dependent on the target
position, and there is a variable amount of error in the end-points
achieved by the model. Colour is used in this diagram as an aid to
distinguishing different saccades and their targets; for a given
saccade, the target location is given by the cross of the same colour
closest to the end of the trajectory. b) The three rotational
components of the `blue' saccade, to target location (-7,-7). c) The
three rotational components of the `red' saccade, to target location
(0,-10).}
\label{fig:outmany}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.8\textwidth]{./figures/lat_vs_everything.png}
\end{center}
\caption{ Exploring saccade latencies. a) Latency to first movement as a
function of target eccentricity for horizontal targets. b) Latency
vs. eccentricity for vertical targets. c) Latency vs. gap at three
different luminance values. The data are shown alongside
the \ccg~model results from Fig.~5 of \cite{cope_basal_2017} in blue,
and the experimental results used in that model in red. The fixation
luminance for the \ccg~curve was 0.5, the target lumianance 0.6 and
the target eccentricity was 8\dg. The difference between
the \ccg~model data and the data from the current model results from
the different mechanism by which activity in SC\_deep causes a
movement, the differing target angle and the reduced fixation
luminance used here. d) The effect
of the dopamine parameter on saccade latencies in gap, step and
overlap conditions, for two different target luminances.  e) Saccade
vs. luminance showing gradual transition between reflexive and express
behaviour.
%f) Saccade duration vs. saccade eccentricity for horizontal
%(rightwards), vertical (downwards) and 45\dg~oblique (down \& right) saccades.
}
\label{fig:latvsall}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.8\textwidth]{./figures/outrtn.png}
\end{center}
\caption{ There and back - a saccade to a target, followed by return to the
original fixation. a) Out and return saccade to a target at (0,-10\dg)
b) Rotational components of the saccade shown in (a). c) Outward and
return trajectories for the saccade shown in (a) alongside saccades to
three other targets.}
\label{fig:outrtn}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.8\textwidth]{./figures/doublesteps.png}
\end{center}
\caption{ Double steps. The effect of illuminating a first target at 8\dg~or
12\dg, followed by a second target at 12\dg~or 8\dg. a) Horizontal
rotation of the eye plotted vs.~time for a saccade to the 12\dg~target
only (red), and to an 8\dg~target at 0.4~s followed by a 12\dg~target
after 30~ms (blue) or 40~ms (black). The timings are indicated by
vertical lines. When the second target is presented up to 30~ms after
the initial target, the initial target has not had time to dominate
the output saccade and a saccade to a location close to the second
target is made. If the delay is 40~ms or more, the activity from the
initial target has time to cause a built up of actiivty in SC\_deep
and an initial saccade close to the first target is made, followed,
after a longer than usual latency period, with a second saccade closer
to the second target. In this graph, the mean of five separate
simulations is plotted along with $\pm$1 standard deviation around the
mean. b) The $\theta_{x}$/$\theta_{y}$ trajectories corresponding to
the data presented in (a).}
\label{fig:doublesteps}
\end{figure}

\end{document}